---
title: "Property Tax Exemption"
author: "Fernando Garcia - Drazzel Feliu"
date: "8/11/2020"
output: html_document
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(lubridate)
library(ggmap)
library(tidycensus)
library(mapview)
library(sf)
library(leaflet)
library(viridis)
library(maps)
library(htmlwidgets)
library(htmltools)
library(tigris)
library(reshape2)
library(ggrepel)
library(ggmap)

rm(list = ls())
```



```{r, warning=FALSE, error=FALSE}
census_data <- get_acs(
  geography = "tract",
  variables = c(
    population_18 = "B01001_001",
    income_18 = "B19013_001",
    agg_income_18 = "B19313_001",
    avg_household_size_18 = "B25010_001",
    value_18 = "B25077_001",
    education_18 = "B15003_022",
    median_rent_18 = "B25064_001",
    median_rent_income_18 = "B25071_001"
  ),
  state = "IL",
  county = "Cook County",
  year = 2018,
  geometry = T
)

```

```{r, warning=FALSE, error=FALSE}
census_data <- census_data %>%
  select(GEOID, NAME, variable, estimate)

census_data <- census_data %>%
  spread(key = variable, value = estimate)

census_data <- census_data %>%
  mutate(bachelor_18=education_18/population_18)

```

Checking summary statistics:

```{r, warning=FALSE, error=FALSE}
census_data %>%
  mutate(income_pc_18=agg_income_18/population_18) %>%
  summarise(mean_income_pc_18=mean(income_pc_18,na.rm=T))

census_data <- census_data %>%
  mutate(income_pc_18=agg_income_18/population_18) #create income per capita

census_data %>%
  summarise(mean_avg_household_size_18 = mean(avg_household_size_18,na.rm=T))

census_data %>%
  mutate(income_household_18=(agg_income_18/population_18) * 
           avg_household_size_18) %>%
  summarise(median_income_household_18=median(income_household_18,na.rm=T))

census_data <- census_data %>%
  mutate(household_income_18=(agg_income_18/population_18) * 
           avg_household_size_18) #create household income

```

Mean income per capita for 2018 is $34,905. Median household income is 76,307 with an average household size of 2.6. Correct if we look at https://www.census.gov/quickfacts/fact/table/chicagocityillinois/LND110210.



#Race data

```{r message=FALSE, warning=FALSE}
codebook <- load_variables(2018,"acs5")

racecodebook <- codebook %>% 
  filter(str_detect(label,"race")) %>% 
  filter(str_detect(label,"Population of one")) %>%
  mutate(names="name", count=1:7, merge=paste(names,count, sep="")) %>%
  select(-c(names,count))

```

```{r message=FALSE, warning=FALSE}
racedataindi <- get_acs(geography = "tract",
        variables = unlist(racecodebook$name),
        state = "IL",
        county = "Cook County",
        geometry = T,
        year = 2018,
        cb = FALSE)
```

```{r message=FALSE, warning=FALSE}

racedata <- racedataindi %>% 
  filter(variable!="C02003_002") %>%
  group_by(GEOID) %>% 
  mutate(sum_18=sum(estimate))

racedata <- merge(racedata,racecodebook[c(1,2)],by.x="variable",by.y="name") %>%
  select(-variable) %>%
  mutate(label=str_remove(label,"Estimate!!Total!!Population of one race!!")) %>% 
  group_by(GEOID) %>% 
  arrange(GEOID)

racedata <- racedata %>%
  select(-c(moe)) %>%
  spread(key = label, value = estimate) %>% 
  arrange(GEOID)

```

We will join both databases:

```{r, warning=FALSE, error=FALSE}

racedata$geometry <- NULL


census_data_18 <- left_join(census_data, racedata, by = c("NAME", "GEOID"))

```

Now, we will pull the dame data for year 2013:

```{r, warning=FALSE, error=FALSE}
census_data <- get_acs(
  geography = "tract",
  variables = c(
    population_13 = "B01001_001",
    income_13 = "B19013_001",
    agg_income_13 = "B19313_001",
    avg_household_size_13 = "B25010_001",
    education_13 = "B15003_022",
    value_13 = "B25077_001",
    median_rent_13 = "B25064_001",
    median_rent_income_13 = "B25071_001"
  ),
  state = "IL",
  county = "Cook County",
  year = 2013,
  geometry = T
)

class(census_data)
```

```{r, warning=FALSE, error=FALSE}
census_data <- census_data %>%
  select(GEOID, NAME, variable, estimate)

census_data <- census_data %>%
  spread(key = variable, value = estimate)

census_data <- census_data %>%
  mutate(bachelor_13=education_13/population_13)

```

```{r, warning=FALSE, error=FALSE}

census_data %>%
  mutate(income_pc_13=agg_income_13/population_13) %>%
  summarise(mean_income_pc_13=mean(income_pc_13,na.rm=T))

census_data <- census_data %>%
  mutate(income_pc_13=agg_income_13/population_13) #create income per capita

census_data %>%
  summarise(mean_avg_household_size_13=mean(avg_household_size_13,na.rm=T))

census_data %>%
  mutate(income_household_13=(agg_income_13/population_13) * 
           avg_household_size_13) %>%
  summarise(median_income_household_13=median(income_household_13,na.rm=T))

census_data <- census_data %>%
  mutate(household_income_13=(agg_income_13/population_13) * 
           avg_household_size_13) #create household income


```

Median household income is \$ 65,390 and per capita income mean is \$29,682 in 2013.

#Race data 2013


```{r message=FALSE, warning=FALSE}

racedataindi <- get_acs(geography = "tract",
        variables = unlist(racecodebook$name),
        state = "IL",
        county = "Cook County",
        geometry = T,
        year = 2013,
        cb = FALSE)
```

```{r message=FALSE, warning=FALSE}

racedata <- racedataindi %>% 
  filter(variable!="C02003_002") %>%
  group_by(GEOID) %>% 
  mutate(sum_13=sum(estimate))

racedata <- merge(racedata,racecodebook[c(1,2)],by.x="variable",by.y="name") %>%
  select(-variable) %>%
  mutate(label=str_remove(label,"Estimate!!Total!!Population of one race!!")) %>% 
  group_by(GEOID) %>% 
  arrange(GEOID)

racedata <- racedata %>%
  select(-c(moe)) %>%
  spread(key = label, value = estimate) %>% 
  arrange(GEOID)

```

We will join both databases for year 2013:

```{r, warning=FALSE, error=FALSE}

racedata$geometry <- NULL

census_data_13 <- left_join(census_data, racedata, by=c ("GEOID", "NAME"))

```

```{r, warning=FALSE, error=FALSE}
census_data_13 <- census_data_13 %>%
  rename("White 13"= "White")

census_data_13 <- census_data_13 %>%
  rename("Some other race 13"="Some other race")

census_data_13 <- census_data_13 %>%
  rename("Native Hawaiian and Other Pacific Islander  13" = 
           "Native Hawaiian and Other Pacific Islander")

census_data_13 <- census_data_13 %>%
  rename("Black or African American  13" = "Black or African American")

census_data_13 <- census_data_13 %>%
  rename("Asian alone  13" = "Asian alone")

census_data_13 <- census_data_13 %>%
  rename("American Indian and Alaska Native  13" = "American Indian and Alaska Native")

```

Then, we will join databases from 2018 and 2013

```{r, warning=FALSE, error=FALSE}
census_data_13$geometry <- NULL

census <- census_data_18 %>%
  left_join(census_data_13, by = c("GEOID", "NAME"))
```


#Restricting to City of Chicago

```{r, warning=FALSE, error=FALSE}

cityboundary <- read_sf("geo_export_72cf3d72-fa82-4448-a382-c9dd6dcb4fd2.shp")

cityboundary <- st_union(cityboundary)

cityboundary <- st_transform(
  cityboundary,
  4326
)

community <- read_sf("geo_export_72cf3d72-fa82-4448-a382-c9dd6dcb4fd2.shp")

community <- st_transform(
  community,
  4326
)

```

```{r, warning=FALSE, error=FALSE}

cityboundary <- st_transform(cityboundary, "+proj=longlat +datum=NAD83 +no_defs")

census <- st_transform(census, "+proj=longlat +datum=NAD83 +no_defs")
```

We do the intersection of both areas:

```{r}
census$inter <- as.numeric(st_intersects(census, cityboundary))

census <- census %>% 
  filter(inter == 1)

census <- census %>%
  select(-c(inter))
```

Filter the areas where the city of Chicago and the cencus tract intersect but not coincide completely:

```{r}

filter_list <- c(
  "17031810000", "17031810200", "17031810301", "17031807600", "17031808002",
  "17031808001", "17031807900", "17031807800", "17031808100", "17031808100",
  "17031805402", "17031805502", "17031810400", "17031805702", "17031770700",
  "17031770602", "17031770500", "17031770800", "17031811600", "17031811701",
  "17031811500", "17031770902", "17031770901", "17031810501", "17031810502",
  "17031810600", "17031811200", "17031811100", "17031810800", "17031810701",
  "17031810702", "17031810900", "17031811900", "17031812200", "17031812100",
  "17031812500", "17031812600", "17031813000", "17031813100", "17031813500",
  "17031813400", "17031813301", "17031813302", "17031813801", "17031814200",
  "17031820700", "17031820800", "17031820300", "17031820400", "17031820502",
  "17031820901", "17031820902", "17031821102", "17031822101", "17031822000",
  "17031821600", "17031821700", "17031821900", "17031822701", "17031822702",
  "17031822802", "17031822801", "17031823200", "17031823303", "17031823304",
  "17031823400", "17031823500", "17031821200", "17031821401", "17031821402",
  "17031821500", "17031826500", "17031826401", "17031825801", "17031825700",
  "17031805600"
)

census <- census %>%
  filter(!GEOID %in% filter_list)
```


Changes between 2018 and 2013

```{r, warning=FALSE, error=FALSE}

census <- census %>%
  mutate(increase_value = (value_18 / value_13 - 1) * 100)

census <- census %>%
  mutate(increase_income = (income_18 / income_13 - 1) * 100)

census <- census %>%
  mutate(increase_pop = (population_18 / population_13 - 1) * 100)

census <- census %>%
  mutate(increase_education = (education_18 / education_13 - 1) * 100)

census <- census %>%
  mutate(increase_bachelor = (bachelor_18 / bachelor_13 - 1) * 100)

```

#Different Maps with increase in property values as example

Map #1

```{r, warning=FALSE, error=FALSE}

mapview(census, zcol = "increase_value")

```

Map #2

```{r, warning=FALSE, error=FALSE}
ggplot() +
  geom_sf(census,
    mapping = aes(
      fill = increase_value,
      color = increase_value
    ),
    inherit.aes = F,
    alpha = .99
  ) +
  scale_fill_viridis() +
  scale_color_viridis() +
  theme_minimal() +
  ggtitle("Percentage increase in property values between 2018 and 2013", )
```

Leaflet Options and Colors:

```{r Leaflet Options - Community Areas}
neigh_h_lights <- highlightOptions(
  weight = 4.5,
  color = "#ffffff",
  fillOpacity = .5,
  bringToFront = TRUE)

neigh_l_options <- labelOptions(
  style = list("font-weight" = "normal", padding = "3px 8px"),
  textsize = "15px",
  direction = "auto")

neigh_labels <- sprintf(
  paste0(
  "<strong>Name: </strong>%s<br/>"),
  community$community
  ) %>% 
  lapply(HTML)
```

```{r Leaflet Options - Census Tracts}
tract_h_lights <- highlightOptions(
  weight = 4.5,
  color = "#ffffff",
  fillOpacity = .5,
  bringToFront = TRUE)

tract_l_options <- labelOptions(
  style = list("font-weight" = "normal", padding = "3px 8px"),
  textsize = "15px",
  direction = "auto")

tract_labels <- sprintf(
  paste0(
  "<strong>Name: </strong>%s<br/>",
  "<strong>Median Income 2013: </strong>%g<br/>",
  "<strong>Median Income 2018: </strong>%g<br/>",
  "<strong>Change in Median Income: </strong>%g<br/>"),
  census$NAME,
  census$income_13,
  census$income_18,
  census$increase_income
  ) %>% 
  lapply(HTML)
```


```{r, warning=FALSE, error=FALSE}

color_200k <- colorNumeric(
  palette = c("white", "blue"),
  domain = census$increase_income
)
```

Map #3:


```{r}

leaflet(options = leafletOptions(zoomSnap = 0.1)) %>%
  addTiles() %>%
  addPolygons(
    data = st_geometry(census),
    color = "black",
    fillColor = color_200k(census$increase_value),
    weight = 2.5,
    fillOpacity = .55,
    smoothFactor = .2
  ) %>%
  addLegend(
    title = "Percentage increase of housing values between 2018 and 2013",
    pal = color_200k,
    values = census$increase_value,
    position = "topright"
  )

```


#Gentrification:

Distribution of home values in 2018:

```{r}

census %>%
  summarise(median_value_18=median(value_18, na.rm=T))

ggplot(census) +
  geom_histogram(aes(x = value_18, y = ..density..),
    binwidth = 10000,
    colour = "black", fill = "white"
  )  +
  geom_vline(xintercept = 226800, color="blue") +
  ggtitle("Histogram price increase in home values")

``` 

We will look into the distribution of the home value increase:


```{r}

census %>%
  summarise(median_value_18=median(value_18, na.rm=T))

ggplot(census) +
  geom_histogram(aes(x = increase_value, y = ..density..),
    binwidth = 1,
    colour = "black", fill = "white"
  ) +
  geom_vline(xintercept = 4.334484, color="blue") +
  ggtitle("Histogram price icnrease home values")

``` 

We will zoom in to look at the outliers:

```{r}
ggplot(census) +
  geom_histogram(aes(x = increase_value, y = ..density..),
    binwidth = 1,
    colour = "black", fill = "white"
  ) +
  xlim(c(25, 200)) +
  ggtitle("Histogram price increase in home values")

``` 

The cumulative distribution looks like this:

```{r, warning=FALSE, error=FALSE}
census_rest <- census %>%
  filter(increase_value > 0)

plot(ecdf(census_rest$increase_value), 
     xlab="Home value price increase ($)",
     ylab="Cumulative Distribution",
     main="Home value price increase distributions from 2013-2018")
abline(h=0.95, col="blue")
```

The last 5% of the distribution gets a 50% or more of an increase in home values. We can use this to define gentrification.

First definition of gentrification:

```{r, warning=FALSE, error=FALSE}
census <- census %>%
  mutate(gentri_1 = case_when(
    increase_value > 50 & value_18 > 200000 ~ 1,
    increase_value > 25 & value_18 > 220000 ~1,
    increase_value <= 50 & (value_18 > 200000 & value_18 < 220000) ~ 0,
    (increase_value > 25 & increase_value <= 50) & value_18 <= 220000 ~ 0,
    increase_value <= 25 ~ 0,
    value_18 <= 200000 ~ 0
    )
  )
```

```{r, warning=FALSE, error=FALSE}

census %>%
  filter(gentri_1==1) %>%
  count()

```

This measure of gentrification reaches 84 tracts.

These are the tracts that would be gentrified according to definition 1:

```{r}

leaflet(options = leafletOptions(zoomSnap = 0.1)) %>%
  addTiles() %>%
  addPolygons(
    data = st_geometry(census %>% filter(census$gentri_1==1)),
    color = "black",
    fillColor = color_200k(census$increase_income),
    weight = 2.5,
    fillOpacity = .3,
    smoothFactor = .2,
    popup = tract_labels,
    popupOptions = tract_l_options,
    highlightOptions = tract_h_lights,
    group = "Gentrification By Education"
  ) %>%
  addPolygons(
    data = st_geometry(community),
    popup = neigh_labels,
    popupOptions = neigh_l_options,
    highlightOptions = neigh_h_lights,
    group = "Community Areas") %>%
  addLegend(
    title = "Gentrified tracts between 2018 and 2013",
    pal = color_200k,
    values = census$increase_value,
    position = "topright"
  ) %>% 
  addLayersControl(
    overlayGroups = c("Gentrification By Education", "Community Areas"),
    position = "bottomright")


```


Second definition of gentrification:

```{r, warning=FALSE, error=FALSE}

census <- census %>%
  mutate(gentri_2 = case_when(
    increase_value > 50 & household_income_13 <= 52000 ~ 1,
    increase_value > 25 & household_income_13 <= 40000 ~1,
    (increase_value > 25 & increase_value <= 50) & 
      (household_income_13 > 40000 & household_income_13 < 52000) ~ 0,
    increase_value <= 25 ~ 0,
    household_income_13 > 52000 ~ 0
    )
  )
```

```{r, warning=FALSE, error=FALSE}

census %>%
  filter(gentri_2==1) %>%
  count()

```

This measure of gentrification reaches 27 tracts.

These are the tracts that would be gentrified according to definition 2:

```{r}

map <- leaflet(options = leafletOptions(zoomSnap = 0.1)) %>%
  addTiles() %>%
  addPolygons(
    data = st_geometry(census %>% filter(census$gentri_2==1)),
    color = "black",
    fillColor = color_200k(census$increase_value),
    weight = 2.5,
    fillOpacity = .55,
    smoothFactor = .2
  ) %>%
  addPolygons(
    data = st_geometry(community)) %>%
  addLegend(
    title = "Gentrified tracts between 2018 and 2013",
    pal = color_200k,
    values = census$increase_value,
    position = "topright"
  )
saveWidget(map, "downstate_200k.html")

map
```

Summary statistics for gentrified areas:


```{r, warning=FALSE, error=FALSE}

census %>%
  filter(gentri_2==1) %>%
  summarise(mean_increase_edu = median(increase_education, na.rm = T))

census %>%
  filter(gentri_2==0) %>%
  summarise(mean_increase_edu = median(increase_education, na.rm = T))

census_gen %>%
  filter(gentri_2==1) %>%
  summarise(mean_increase_bc = median(increase_bachelor, na.rm = T))

census %>%
  filter(gentri_1==0) %>%
  summarise(mean_increase_bc = median(increase_bachelor, na.rm = T))

census %>%
  filter(gentri_2==1) %>%
  summarise(mean_increase_income = mean(increase_income, na.rm = T))

census %>%
  filter(gentri_2==0) %>%
  summarise(mean_increase_income = mean(increase_income, na.rm = T))


```
