---
title: "Property Tax Exemption"
author: "Fernando Garcia - Drazzel Feliu"
date: "8/11/2020"
output: html_document
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(lubridate)
library(ggmap)
library(tidycensus)
library(mapview)
library(sf)
library(leaflet)
library(viridis)
library(maps)
library(htmlwidgets)
library(htmltools)
library(tigris)
library(reshape2)
library(ggrepel)
library(ggmap)
```


```{r, warning=FALSE, error=FALSE}
census_data <- get_acs(
  geography = "tract",
  variables = c(
    population_18 = "B01001_001",
    income_18 = "B19013_001",
    value_18 = "B25077_001"
  ),
  state = "IL",
  county = "Cook County",
  year = 2018,
  geometry = T
)

```

```{r, warning=FALSE, error=FALSE}
census_data <- census_data %>%
  select(GEOID, NAME, variable, estimate)

census_data <- census_data %>%
  spread(key = variable, value = estimate)

```

#Race data

```{r message=FALSE, warning=FALSE}
codebook <- load_variables(2017,"acs5")

racecodebook <- codebook %>% 
  filter(str_detect(label,"race")) %>% 
  filter(str_detect(label,"Population of one")) %>%
  mutate(names="name", count=1:7, merge=paste(names,count, sep="")) %>%
  select(-c(names,count))

```

```{r message=FALSE, warning=FALSE}
racedataindi <- get_acs(geography = "tract",
        variables = unlist(racecodebook$name),
        state = "IL",
        county = "Cook County",
        geometry = T,
        year = 2018,
        cb = FALSE)
```

```{r message=FALSE, warning=FALSE}

racedata <- racedataindi %>% 
  filter(variable!="C02003_002") %>%
  group_by(GEOID) %>% 
  mutate(sum_18=sum(estimate))

racedata <- merge(racedata,racecodebook[c(1,2)],by.x="variable",by.y="name") %>%
  select(-variable) %>%
  mutate(label=str_remove(label,"Estimate!!Total!!Population of one race!!")) %>% 
  group_by(GEOID) %>% 
  arrange(GEOID)

racedata <- racedata %>%
  select(-c(moe)) %>%
  spread(key = label, value = estimate) %>% 
  arrange(GEOID)

```

We will join both databases:

```{r, warning=FALSE, error=FALSE}

racedata$geometry <- NULL


census_data_18 <- left_join(census_data, racedata, by = c("NAME", "GEOID"))

```

Now, we will pull the dame data for year 2013:

```{r, warning=FALSE, error=FALSE}
census_data <- get_acs(
  geography = "tract",
  variables = c(
    population_13 = "B01001_001",
    income_13 = "B19013_001",
    value_13 = "B25077_001"
  ),
  state = "IL",
  county = "Cook County",
  year = 2013,
  geometry = T
)

class(census_data)
```

```{r, warning=FALSE, error=FALSE}
census_data <- census_data %>%
  select(GEOID, NAME, variable, estimate)

census_data <- census_data %>%
  spread(key = variable, value = estimate)

```

#Race data 2013


```{r message=FALSE, warning=FALSE}

racedataindi <- get_acs(geography = "tract",
        variables = unlist(racecodebook$name),
        state = "IL",
        county = "Cook County",
        geometry = T,
        year = 2013,
        cb = FALSE)
```

```{r message=FALSE, warning=FALSE}

racedata <- racedataindi %>% 
  filter(variable!="C02003_002") %>%
  group_by(GEOID) %>% 
  mutate(sum_13=sum(estimate))

racedata <- merge(racedata,racecodebook[c(1,2)],by.x="variable",by.y="name") %>%
  select(-variable) %>%
  mutate(label=str_remove(label,"Estimate!!Total!!Population of one race!!")) %>% 
  group_by(GEOID) %>% 
  arrange(GEOID)

racedata <- racedata %>%
  select(-c(moe)) %>%
  spread(key = label, value = estimate) %>% 
  arrange(GEOID)

```

We will join both databases for year 2013:

```{r, warning=FALSE, error=FALSE}

racedata$geometry <- NULL

census_data_13 <- left_join(census_data, racedata, by=c ("GEOID", "NAME"))

```

```{r, warning=FALSE, error=FALSE}
census_data_13 <- census_data_13 %>%
  rename("White 13"= "White")

census_data_13 <- census_data_13 %>%
  rename("Some other race 13"="Some other race")

census_data_13 <- census_data_13 %>%
  rename("Native Hawaiian and Other Pacific Islander  13" = 
           "Native Hawaiian and Other Pacific Islander")

census_data_13 <- census_data_13 %>%
  rename("Black or African American  13" = "Black or African American")

census_data_13 <- census_data_13 %>%
  rename("Asian alone  13" = "Asian alone")

census_data_13 <- census_data_13 %>%
  rename("American Indian and Alaska Native  13" = "American Indian and Alaska Native")

```

Then, we will join databases from 2018 and 2013

```{r, warning=FALSE, error=FALSE}
census_data_13$geometry <- NULL

census <- census_data_18 %>%
  left_join(census_data_13, by = c("GEOID", "NAME"))
```


#Restricting to City of Chicago

```{r, warning=FALSE, error=FALSE}

cityboundary <- read_sf("geo_export_72cf3d72-fa82-4448-a382-c9dd6dcb4fd2.shp")

cityboundary <- st_union(cityboundary)

cityboundary <- st_transform(
  cityboundary,
  4326
)
```

```{r, warning=FALSE, error=FALSE}

cityboundary <- st_transform(cityboundary, "+proj=longlat +datum=NAD83 +no_defs")

census <- st_transform(census, "+proj=longlat +datum=NAD83 +no_defs")
```

We do the intersection of both areas:

```{r}
census$inter <- as.numeric(st_intersects(census, cityboundary))

census <- census %>% 
  filter(inter == 1)
```

Filter the areas where the city of Chicago and the cencus tract intersect but not coincide completely:

```{r}

filter_list <- c(
  "17031810000", "17031810200", "17031810301", "17031807600", "17031808002",
  "17031808001", "17031807900", "17031807800", "17031808100", "17031808100",
  "17031805402", "17031805502", "17031810400", "17031805702", "17031770700",
  "17031770602", "17031770500", "17031770800", "17031811600", "17031811701",
  "17031811500", "17031770902", "17031770901", "17031810501", "17031810502",
  "17031810600", "17031811200", "17031811100", "17031810800", "17031810701",
  "17031810702", "17031810900", "17031811900", "17031812200", "17031812100",
  "17031812500", "17031812600", "17031813000", "17031813100", "17031813500",
  "17031813400", "17031813301", "17031813302", "17031813801", "17031814200",
  "17031820700", "17031820800", "17031820300", "17031820400", "17031820502",
  "17031820901", "17031820902", "17031821102", "17031822101", "17031822000",
  "17031821600", "17031821700", "17031821900", "17031822701", "17031822702",
  "17031822802", "17031822801", "17031823200", "17031823303", "17031823304",
  "17031823400", "17031823500", "17031821200", "17031821401", "17031821402",
  "17031821500", "17031826500", "17031826401", "17031825801", "17031825700",
  "17031805600"
)

census <- census %>%
  filter(!GEOID %in% filter_list)
```

Checking the limits:

```{r, warning=FALSE, error=FALSE}

leaflet() %>%
  addProviderTiles(providers$OpenStreetMap) %>%
  addPolygons(data = st_geometry(census))

```


Changes between 2018 and 2013

```{r, warning=FALSE, error=FALSE}

census <- census %>%
  mutate(increase_value = (value_18 / value_13 - 1) * 100)

census <- census %>%
  mutate(increase_income = (income_18 / income_13 - 1) * 100)

census <- census %>%
  mutate(increase_pop = (population_18 / population_13 - 1) * 100)


```

#Different Maps with increase in property values as example

Map #1

```{r, warning=FALSE, error=FALSE}

mapview(census, zcol = "increase_value")

```

Map #2

```{r, warning=FALSE, error=FALSE}
ggplot() +
  geom_sf(census,
    mapping = aes(
      fill = increase_value,
      color = increase_value
    ),
    inherit.aes = F,
    alpha = .99
  ) +
  scale_fill_viridis() +
  scale_color_viridis() +
  theme_minimal() +
  ggtitle("Percentage increase in property houses between 2018 and 2013", )
```

Colors:

```{r}

# calabels <- sprintf(
# paste0("<strong>%g%%</strong> of single filers earning over $200K in Illinois reside in <strong>%s.</strong><br/>"),
#  round(test$x_200_k_or_more_as_pct,2),
# test$county) %>%
# lapply(HTML)

# callaboptions <- labelOptions(
# style = list("font-weight" = "normal",
#             padding = "3px 8px"),
# textsize = "15px",
# direction = "auto")
```

```{r, warning=FALSE, error=FALSE}

color_200k <- colorNumeric(
  palette = c("white", "blue"),
  domain = chicago_only$increase_value
)
```

Map #3:


```{r}

map <- leaflet(options = leafletOptions(zoomSnap = 0.1)) %>%
  addTiles() %>%
  addPolygons(
    data = st_geometry(census),
    color = "black",
    fillColor = color_200k(census$increase_value),
    weight = 2.5,
    fillOpacity = .55,
    smoothFactor = .2
  ) %>%
  addLegend(
    title = "Percentage increase of housing values between 2018 and 2013",
    pal = color_200k,
    values = census$increase_value,
    position = "topright"
  )
saveWidget(map, "downstate_200k.html")

map
```
