---
title: "Property Tax Exemption"
author: "Fernando Garcia - Drazzel Feliu"
date: "8/11/2020"
output: html_document
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(lubridate)
library(ggmap)
library(tidycensus)
library(mapview)
library(sf)
library(leaflet)
library(viridis)
library(maps)
library(htmlwidgets)
library(htmltools)
library(tigris)
library(reshape2)
library(ggrepel)
library(ggmap)

rm(list = ls())
```

We pull data from the ACS using tidycensus:

```{r, warning=FALSE, error=FALSE}
census_data <- get_acs(
  geography = "tract",
  variables = c(
    population_18 = "B01001_001",
    income_18 = "B19013_001", # median household income
    agg_income_18 = "B19313_001",
    avg_household_size_18 = "B25010_001",
    value_18 = "B25077_001", # median home values
    bachelor_18 = "B15003_022", # population with a bachelor's degree
    master_18 = "B15003_023", # population with a master's degree
    prof_18 = "B15003_024", # population with a professional's degree
    phd_18 = "B15003_025", # population with a phd
    median_rent_18 = "B25064_001",
    median_rent_income_18 = "B25071_001", # median rent over income
    median_year_built = "B25035_001",
    year_built_18 = "B25034_001",
    built_2014_18 = "B25034_002"
  ),
  state = "IL",
  county = "Cook County",
  year = 2018,
  geometry = T
)
```

```{r, warning=FALSE, error=FALSE}
census_data <- census_data %>%
  select(GEOID, NAME, variable, estimate)

census_data <- census_data %>%
  spread(key = variable, value = estimate)
```

We create the variable people with bachelor degree or more:

```{r, warning=FALSE, error=FALSE}
census_data <- census_data %>%
  mutate(bc_18 = bachelor_18 + master_18 + prof_18 + phd_18)

census_data <- census_data %>%
  select(-c(bachelor_18, master_18, prof_18, phd_18))
```

We create income per capita and household income for year 2018:

```{r, warning=FALSE, error=FALSE}

census_data <- census_data %>%
  mutate(income_pc_18 = agg_income_18 / population_18) # create income per capita


census_data <- census_data %>%
  mutate(household_income_18 = (agg_income_18 / population_18) *
    avg_household_size_18) # create household income
```

#Race data

```{r message=FALSE, warning=FALSE}
codebook <- load_variables(2018, "acs5")

racecodebook <- codebook %>%
  filter(str_detect(label, "race")) %>%
  filter(str_detect(label, "Population of one race")) %>%
  mutate(names = "name", count = 1:7, merge = paste(names, count, sep = "")) %>%
  select(-c(names, count))
```

```{r message=FALSE, warning=FALSE}
racedataindi <- get_acs(
  geography = "tract",
  variables = c(
    unlist(racecodebook$name),
    "C02003_009",
    "B03003_003",
    "B03002_003",
    "B03002_004"
  ),
  state = "IL",
  county = "Cook County",
  geometry = T,
  year = 2018,
  cb = FALSE
)
```

```{r message=FALSE, warning=FALSE}

racedata <- racedataindi %>%
  filter(variable != "C02003_002") %>%
  group_by(GEOID) %>%
  mutate(sum_18 = sum(estimate))

racedata <- racedata %>%
  select(-c(moe)) %>%
  spread(key = variable, value = estimate) %>%
  arrange(GEOID)

racedata <- racedata %>%
  rename(
    white_18 = C02003_003,
    black_18 = C02003_004,
    indian_native_18 = C02003_005,
    asian_18 = C02003_006,
    native_pacific_18 = C02003_007,
    other_race_18 = C02003_008,
    two_or_more_18 = C02003_009,
    latino_18 = B03003_003,
    white_not_latino_18 = B03002_003,
    black_not_latino_18 = B03002_004
  )

racedata <- racedata %>%
  mutate(sum_18 = sum_18 - latino_18 - white_not_latino_18 - black_not_latino_18)
```

We will join both databases for 2018:

```{r, warning=FALSE, error=FALSE}

racedata$geometry <- NULL


census_data_18 <- left_join(census_data, racedata, by = c("NAME", "GEOID"))
```

#Year structure built 2018

```{r message=FALSE, warning=FALSE}

builtcodebook <- codebook %>%
  filter(str_detect(concept, "^YEAR STRUCTURE BUILT$")) %>%
  filter(str_detect(label, "Built")) %>%
  mutate(names = "name", count = 1:10, merge = paste(names, count, sep = "")) %>%
  select(-c(names, count))
```

```{r message=FALSE, warning=FALSE}
builtdataindi <- get_acs(
  geography = "tract",
  variables = unlist(builtcodebook$name),
  state = "IL",
  county = "Cook County",
  geometry = T,
  year = 2018,
  cb = FALSE
)
```

```{r message=FALSE, warning=FALSE}

builtdata <- builtdataindi %>%
  select(-c(moe)) %>%
  spread(key = variable, value = estimate) %>%
  arrange(GEOID)

builtdata <- builtdata %>%
  rename(
    built_14_18 = B25034_002, #structure built from 2014 to 2018
    built_10_13_18 = B25034_003, #structure built from 2010 to 2012
    built_00_09_18 = B25034_004, #structure built from 2000 to 2009
    built_90_99_18 = B25034_005, #structure built from 1990 to 1999
    built_80_89_18 = B25034_006, #structure built from 1980 to 1989
    built_70_79_18 = B25034_007, #structure built from 1970 to 1979
    built_60_69_18 = B25034_008, #structure built from 1960 to 1969
    built_50_59_18 = B25034_009, #structure built from 1950 to 1959
    built_40_49_18 = B25034_010, #structure built from 1940 to 1949
    built_39_18 = B25034_011 #structure built 1939 or before
  )
```

We will join both databases for 2018:

```{r, warning=FALSE, error=FALSE}

builtdata$geometry <- NULL

census_data_18 <- left_join(census_data_18, builtdata, by = c("NAME", "GEOID"))
```


Now, we will pull the same data for year 2013:

```{r, warning=FALSE, error=FALSE}
census_data <- get_acs(
  geography = "tract",
  variables = c(
    population_13 = "B01001_001",
    income_13 = "B19013_001",
    agg_income_13 = "B19313_001",
    avg_household_size_13 = "B25010_001",
    bachelor_13 = "B15003_022",
    master_13 = "B15003_023",
    prof_13 = "B15003_024",
    phd_13 = "B15003_025",
    value_13 = "B25077_001",
    median_rent_13 = "B25064_001",
    median_rent_income_13 = "B25071_001"
  ),
  state = "IL",
  county = "Cook County",
  year = 2013,
  geometry = T
)

class(census_data)
```

```{r, warning=FALSE, error=FALSE}
census_data <- census_data %>%
  select(GEOID, NAME, variable, estimate)

census_data <- census_data %>%
  spread(key = variable, value = estimate)

```

```{r, warning=FALSE, error=FALSE}
census_data <- census_data %>%
  mutate(bc_13 = bachelor_13 + master_13 + prof_13 + phd_13)

census_data <- census_data %>%
  select(-c(bachelor_13, master_13, prof_13, phd_13))
```

```{r, warning=FALSE, error=FALSE}

census_data <- census_data %>%
  mutate(income_pc_13 = agg_income_13 / population_13) # create income per capita


census_data <- census_data %>%
  mutate(household_income_13 = (agg_income_13 / population_13) *
    avg_household_size_13) # create household income
```


#Race data 2013


```{r message=FALSE, warning=FALSE}
racedataindi <- get_acs(
  geography = "tract",
  variables = c(
    unlist(racecodebook$name),
    "C02003_009",
    "B03003_003",
    "B03002_003",
    "B03002_004"
  ),
  state = "IL",
  county = "Cook County",
  geometry = T,
  year = 2013,
  cb = FALSE
)
```

```{r message=FALSE, warning=FALSE}

racedata <- racedataindi %>%
  filter(variable != "C02003_002") %>%
  group_by(GEOID) %>%
  mutate(sum_13 = sum(estimate))

racedata <- racedata %>%
  select(-c(moe)) %>%
  spread(key = variable, value = estimate) %>%
  arrange(GEOID)

racedata <- racedata %>%
  rename(
    white_13 = C02003_003,
    black_13 = C02003_004,
    indian_native_13 = C02003_005,
    asian_13 = C02003_006,
    native_pacific_13 = C02003_007,
    other_race_13 = C02003_008,
    two_or_more_13 = C02003_009,
    latino_13 = B03003_003,
    white_not_latino_13 = B03002_003,
    black_not_latino_13 = B03002_004
  )
racedata <- racedata %>%
  mutate(sum_13 = sum_13 - latino_13 - white_not_latino_13 - black_not_latino_13)
```

We will join both databases for year 2013:

```{r, warning=FALSE, error=FALSE}

racedata$geometry <- NULL

census_data_13 <- left_join(census_data, racedata, by = c("GEOID", "NAME"))
```

#Year structure built 2013

```{r message=FALSE, warning=FALSE}
codebook <- load_variables(2013, "acs5")


builtcodebook <- codebook %>%
  filter(str_detect(concept, "^YEAR STRUCTURE BUILT$")) %>%
  filter(str_detect(label, "Built")) %>%
  mutate(names = "name", count = 1:9, merge = paste(names, count, sep = "")) %>%
  select(-c(names, count))
```

```{r message=FALSE, warning=FALSE}
builtdataindi <- get_acs(
  geography = "tract",
  variables = unlist(builtcodebook$name),
  state = "IL",
  county = "Cook County",
  geometry = T,
  year = 2013,
  cb = FALSE
)
```

```{r message=FALSE, warning=FALSE}

builtdata <- builtdataindi %>%
  select(-c(moe)) %>%
  spread(key = variable, value = estimate) %>%
  arrange(GEOID)

builtdata <- builtdata %>%
  rename(
    built_10_13 = B25034_002,
    built_00_09_13 = B25034_003,
    built_90_99_13 = B25034_004,
    built_80_89_13 = B25034_005,
    built_70_79_13 = B25034_006,
    built_60_69_13 = B25034_007,
    built_50_59_13 = B25034_008,
    built_40_49_13 = B25034_009,
    built_39_13 = B25034_010
  )
```

We will join both databases for 2018:

```{r, warning=FALSE, error=FALSE}

builtdata$geometry <- NULL

census_data_13 <- left_join(census_data_13, builtdata, by = c("NAME", "GEOID"))
```


Then, we will join databases from 2018 and 2013

```{r, warning=FALSE, error=FALSE}
census_data_13$geometry <- NULL

census <- census_data_18 %>%
  left_join(census_data_13, by = c("GEOID", "NAME"))
```


#Restricting to City of Chicago

```{r, warning=FALSE, error=FALSE}

cityboundary <- read_sf("geo_export_72cf3d72-fa82-4448-a382-c9dd6dcb4fd2.shp")

cityboundary <- st_union(cityboundary)

cityboundary <- st_transform(
  cityboundary,
  4326
)

community <- read_sf("geo_export_72cf3d72-fa82-4448-a382-c9dd6dcb4fd2.shp")

community <- st_transform(
  community,
  4326
)
```

```{r, warning=FALSE, error=FALSE}

cityboundary <- st_transform(cityboundary, "+proj=longlat +datum=NAD83 +no_defs")

census <- st_transform(census, "+proj=longlat +datum=NAD83 +no_defs")
```

We do the intersection of both areas:

```{r}
census$inter <- as.numeric(st_intersects(census, cityboundary))

census <- census %>%
  filter(inter == 1)

census <- census %>%
  select(-c(inter))
```

Filter the areas where the city of Chicago and the cencus tract intersect but not coincide completely:

```{r}

filter_list <- c(
  "17031810000", "17031810200", "17031810301", "17031807600", "17031808002",
  "17031808001", "17031807900", "17031807800", "17031808100", "17031808100",
  "17031805402", "17031805502", "17031810400", "17031805702", "17031770700",
  "17031770602", "17031770500", "17031770800", "17031811600", "17031811701",
  "17031811500", "17031770902", "17031770901", "17031810501", "17031810502",
  "17031810600", "17031811200", "17031811100", "17031810800", "17031810701",
  "17031810702", "17031810900", "17031811900", "17031812200", "17031812100",
  "17031812500", "17031812600", "17031813000", "17031813100", "17031813500",
  "17031813400", "17031813301", "17031813302", "17031813801", "17031814200",
  "17031820700", "17031820800", "17031820300", "17031820400", "17031820502",
  "17031820901", "17031820902", "17031821102", "17031822101", "17031822000",
  "17031821600", "17031821700", "17031821900", "17031822701", "17031822702",
  "17031822802", "17031822801", "17031823200", "17031823303", "17031823304",
  "17031823400", "17031823500", "17031821200", "17031821401", "17031821402",
  "17031821500", "17031826500", "17031826401", "17031825801", "17031825700",
  "17031805600"
)

census <- census %>%
  filter(!GEOID %in% filter_list)
```

We will look into NAs: 

```{r}
# sapply(census, function(x) sum(is.na(x)))
```

Most of the variables have 3 NAs. These NAs are from O'Hare airport, Midway airport, and Taylor Park (Census Tract 3817) with no population.

Share in race and hispanic population 

```{r, warning=FALSE, error=FALSE}

census <- census %>%
  mutate(share_white_18 = white_18 / sum_18,  #share of white people 2018
         share_black_18 = black_18 / sum_18, #share of black people 2018
         share_latino_18 = latino_18 / sum_18, #share of latino people 2018
         share_white_only_18 = white_not_latino_18 / sum_18,  #share of white not latino people 2018
         share_white_13 = white_13 / sum_13, #share of white people 2013
         share_black_13 = black_13 / sum_13, #share of black people 2013
         share_latino_13 = latino_13 / sum_13, #share of latino people 2013
         share_white_only_13 = white_not_latino_13 / sum_13 #share of white not latino people 2013
         )

```

Outcome changes between 2018 and 2013:

```{r, warning=FALSE, error=FALSE}

census <- census %>%
  mutate(increase_value = (value_18 / value_13 - 1) * 100,
         increase_income = (household_income_18 / household_income_13 - 1) * 100,
         increase_income_pc = (income_pc_18 / income_pc_13 - 1) * 100,
         increase_pop = (population_18 / population_13 - 1) * 100,
         increase_rent = (median_rent_18 / median_rent_13 - 1) * 100,
         increase_share_latino = share_latino_18 / share_latino_13,
         increase_share_black = share_black_18 / share_black_13
         )

is.na(census$increase_share_latino) <- sapply(census$increase_share_latino, is.infinite) #changing the infinite to NAs

is.na(census$increase_share_black) <- 
  sapply(census$increase_share_black, is.infinite)  #changing the infinite to NAs

census$increase_share_black[is.nan(census$increase_share_black)] <- NA #changing the NaN to NA

census$increase_share_latino[is.nan(census$increase_share_latino)] <- NA

```


What limit should we impose on population?

```{r, warning=FALSE, error=FALSE}

census %>%
  summarise(median_pop_18 = median(population_18, na.rm = T))

ggplot(census) +
  geom_histogram(aes(x = population_18, y = ..density..),
    binwidth = 100,
    colour = "black", fill = "white"
  ) +
  geom_vline(xintercept = 3069.5, color = "blue") +
  ggtitle("Histogram of Population by tracts")
```

```{r}
ggplot(census) +
  geom_histogram(aes(x = population_18, y = ..density..),
    binwidth = 50,
    colour = "black", fill = "white"
  ) +
  xlim(c(0, 3500)) +
  ggtitle("Histogram price increase in home values")
``` 


```{r, warning=FALSE, error=FALSE}
population <- census$population_18

quantile(population, c(.01, .05, .10, .15, .20), na.rm = T)
```

Looking at the percentiles we can see that the distribution accumulates more observations at the 10th percentile. We can check how the mean per capita income stabilize after we restrict the sample to these thresholds: 

```{r, warning=FALSE, error=FALSE}

census %>%
  filter(population_18 < 773) %>%
  summarise(mean_income_pc = mean(income_pc_18, na.rm = T))
```

```{r, warning=FALSE, error=FALSE}

census %>%
  filter(population_18 < 1129.75) %>%
  summarise(mean_income_pc = mean(income_pc_18, na.rm = T))
```

```{r, warning=FALSE, error=FALSE}

census %>%
  filter(population_18 < 1425.50) %>%
  summarise(mean_income_pc = mean(increase_income_pc, na.rm = T))
```

```{r, warning=FALSE, error=FALSE}

census %>%
  filter(population_18 < 1592.50) %>%
  summarise(mean_income_pc = mean(increase_income_pc, na.rm = T))
```

```{r, warning=FALSE, error=FALSE}

census %>%
  filter(population_18 < 1756.00) %>%
  summarise(mean_income_pc = mean(increase_income_pc, na.rm = T))
```

```{r, warning=FALSE, error=FALSE}

census %>%
  summarise(mean_income_pc = mean(increase_income_pc, na.rm = T))
```

The mean per capita income stabilize around the 10th percentile of the population in 2018.

Is there a correlation between census tract's size and income?

```{r, warning=FALSE, error=FALSE}

ggplot(data = census) +
  geom_point(mapping = aes(x = income_18, y = population_18)) +
  geom_smooth(mapping = aes(x = income_18, y = population_18), method = lm)
```

```{r, warning=FALSE, error=FALSE}

census %>%
  filter(population_18 > 1425.50) %>%
  ggplot() +
  geom_point(mapping = aes(x = income_18, y = population_18)) +
  geom_smooth(mapping = aes(x = income_18, y = population_18), method = lm)
```

We will restrict tracts with a population of 1426 or more. 

```{r, warning=FALSE, error=FALSE}

census <- census %>%
  filter(population_18 > 1425.50)
```

```{r, warning=FALSE, error=FALSE}
saveRDS(census, file = "census.RDS")
```
