---
title: "Statistics on Gentrification"
author: "Fernando Garcia - Drazzel Feliu"
date: "8/25/2020"
output: html_document
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(lubridate)
library(ggmap)
library(tidycensus)
library(mapview)
library(sf)
library(leaflet)
library(viridis)
library(maps)
library(htmlwidgets)
library(htmltools)
library(tigris)
library(reshape2)
library(ggrepel)
library(ggmap)
library(tables)
library(qwraps2)


rm(list = ls())
```

```{r, warning=FALSE, error=FALSE}

census <- readRDS("census.RDS")
```

Summary statistics overall in levels:

```{r, warning=FALSE, error=FALSE}

census_group_18 <- census %>%
  select(
    household_income_18,
    income_pc_18,
    value_18,
    median_rent_18,
    share_white_only_18,
    share_black_not_latino_18,
    share_latino_18,
    population_18
  )

# functions for mean, sd, min, median, max

Mean <- function(x) base::mean(x, na.rm = TRUE)

Median <- function(x) {
  n <- length(x[!is.na(x)])
  s <- sort(x[!is.na(x)])
  ifelse(n %% 2 == 1, s[(n + 1) / 2], mean(s[n / 2 + 0:1]))
}

std <- function(x) {
  n <- length(x[!is.na(x)])
  m <- var(x, na.rm = T)
  sqrt((m / (n - 1)))
}


Min <- function(x) {
  min(x, na.rm = T)
}

Max <- function(x) {
  max(x, na.rm = T)
}


table_levels_18 <- tabular(Format(digits = 0) * household_income_18 + Format(digits = 0) * income_pc_18 + Format(digits = 0) * value_18 + Format(digits = 0) * median_rent_18 + Format(digits = 0) * population_18 + Format(digits = 1) * share_white_only_18 + Format(digits = 1) * share_black_not_latino_18 + Format(digits = 1) * share_latino_18 ~ (n = 1) + Justify(c, r) * (Mean + std + Min + Median + Max), data = census_group_18)


options("scipen" = 100, "digits" = 0)

toLatex(table_levels_18)
```

```{r, warning=FALSE, error=FALSE}

census_group_13 <- census %>%
  select(
    household_income_13,
    income_pc_13,
    value_13,
    median_rent_13,
    share_white_only_13,
    share_black_not_latino_13,
    share_latino_13,
    population_13
  )

table_levels_13 <- tabular(Format(digits = 0) * household_income_13 + Format(digits = 0) * income_pc_13 + Format(digits = 0) * value_13 + Format(digits = 0) * median_rent_13 + Format(digits = 0) * population_13 + Format(digits = 2) * share_white_only_13 + Format(digits = 2) * share_black_not_latino_13 + Format(digits = 2) * share_latino_13 ~ (n <- 1) + Justify(c, r) * (Mean + std + Min + Median + Max), data = census_group_13)

toLatex(table_levels_13)
```


Mean income per capita for 2018 is $33,599. Median household income is \$68,000 with an average household size of 2.6. Correct if we look at https://www.census.gov/quickfacts/fact/table/chicagocityillinois/LND110210.

Summary statistics for 2013:



Median household income is \$ 58,676.28 and per capita income mean is \$28,747.67 in 2013.


#Gentrification:

Distribution of home values in 2018:

```{r}

census %>%
  summarise(median_value_18 = median(value_18, na.rm = T))

ggplot(census) +
  geom_histogram(aes(x = value_18, y = ..density..),
    binwidth = 10000,
    colour = "black", fill = "white"
  ) +
  geom_vline(xintercept = 236600, color = "blue") +
  ggtitle("Histogram price increase in home values")
``` 

Distribution of home value increase:

```{r}

# census %>%
# summarise(median_value_increase=median(increase_value, na.rm=T))

ggplot(census) +
  geom_histogram(aes(x = increase_value, y = ..density..),
    binwidth = 0.01,
    colour = "black", fill = "white"
  ) +
  geom_vline(xintercept = 0.04437273, color = "blue") +
  ggtitle("Histogram price increase home values")
``` 

We will zoom in to look at the outliers:

```{r}
ggplot(census) +
  geom_histogram(aes(x = increase_value, y = ..density..),
    binwidth = 0.01,
    colour = "black", fill = "white"
  ) +
  xlim(c(0.25, 2)) +
  ggtitle("Histogram price increase in home values")
``` 

```{r, warning=FALSE, error=FALSE}
census_rest <- census %>%
  filter(increase_value > 0)

plot(ecdf(census_rest$increase_value),
  xlab = "Home value price increase ($)",
  ylab = "Cumulative Distribution",
  main = "Home value price increase distributions from 2013-2018"
)
abline(h = 0.95, col = "blue")
```


The cumulative distribution looks like this:

```{r, warning=FALSE, error=FALSE}
homeincrease <- census$increase_value

quantile(homeincrease, c(.5, .75, .8, .85, .9, .95), na.rm = T)
```

The last 10% of the distribution gets a 36% or more of an increase in home values. We can use this to define gentrification.


#Gentrification definitions according to Brummet, Reed, 2019. 

```{r, warning=FALSE, error=FALSE}
census <- census %>%
  mutate(gentri_3 = (bc_18 - bc_13) / population_13)
# change in bachelor degree or higher normalizing by population in 2013
```

```{r}

# census %>%
# summarise(median_gentri_3 = median(gentri_3, na.rm = T))

ggplot(census) +
  geom_histogram(aes(x = gentri_3, y = ..density..),
    colour = "blue", fill = "white",
    bins = 80
  ) +
  geom_vline(xintercept = quantile(gentrification, c(.75), na.rm = T), color = "blue", linetype="dashed") +
  geom_vline(xintercept = quantile(gentrification, c(.90), na.rm = T), color = "red", linetype="dashed") +
  geom_text(aes(0.075, 10,
    label = "75th",
    hjust = 0.1
  )) +
  geom_text(aes(0.125, 10,
    label = "90th",
    hjust = 0.1
  ) ) +
  ggtitle("Histogram Gentrification #1 (Education)") + 
  xlab("Gentrification definition 1")
``` 

What is the th percentile of the gentrification's indicator:

```{r, warning=FALSE, error=FALSE}
gentrification <- census$gentri_3

quantile(gentrification, c(.5, .75, .8, .85, .90, .95), na.rm = T)
```

#Defitinion with the 75th percentile threshold

We create a dummy variable if gentrification is beyond the 75th percentile of this indicator:

```{r, warning=FALSE, error=FALSE}

# census %>%
# summarise(median_income_13=median(household_income_13, na.rm = T))

census <- census %>%
  mutate(one_gentri_1 = case_when(
    gentri_3 >= 0.06704281 & household_income_13 < 58676 ~ 1,
    household_income_13 > 58676 ~ 0,
    gentri_3 < 0.06704281 ~ 0
  ))
```

```{r, warning=FALSE, error=FALSE}
census %>%
  filter(one_gentri_1 == 1) %>%
  count()
```

This definition reaches 59 tracts.


#Defitinion with the 90th percentile threshold

We create a dummy variable if gentrification is beyond the 90th percentile of this indicator:

```{r, warning=FALSE, error=FALSE}

census <- census %>%
  mutate(one_gentri_2 = case_when(
    gentri_3 >= 0.11469142 & household_income_13 < 58676 ~ 1,
    household_income_13 > 58676 ~ 0,
    gentri_3 < 0.11469142 ~ 0
  ))
```

```{r, warning=FALSE, error=FALSE}
census %>%
  filter(one_gentri_2 == 1) %>%
  count()
```

This measure of gentrification reaches 22 tracts.


Gentrification definition number 4 using racial data:

```{r, warning=FALSE, error=FALSE}

census <- census %>%
  mutate(gentri_4 = (white_not_latino_18 - white_not_latino_13) / sum_13)
# change in white not latino population normalizing by population in 2013
```


```{r}

# census %>%
# summarise(median_gentri_4=median(gentri_4, na.rm=T))

ggplot(census) +
  geom_histogram(aes(x = gentri_4, y = ..density..),
    colour = "black", fill = "white"
  ) +
  geom_vline(xintercept = 0.002240093, color = "blue") +
  ggtitle("Histogram price increase in home values")
``` 

What is the th percentile of the gentrification's indicator:

```{r, warning=FALSE, error=FALSE}
gentrification <- census$gentri_4

quantile(gentrification, c(.5, .75, .8, .85, .90, .95), na.rm = T)
```

#Defitinion with the 75th percentile threshold

We create a dummy variable if gentrification is beyond the 75th percentile of this indicator:

```{r, warning=FALSE, error=FALSE}

# census %>%
# summarise(median_income_13=median(household_income_13, na.rm = T))

census <- census %>%
  mutate(one_gentri_3 = case_when(
    gentri_4 >= 0.038203177 & household_income_13 < 58676 ~ 1,
    household_income_13 > 58676 ~ 0,
    gentri_4 < 0.038203177 ~ 0
  ))
```

```{r, warning=FALSE, error=FALSE}
census %>%
  filter(one_gentri_3 == 1) %>%
  count()
```

Definition number 3 reaches 64 tracts.

#Defitinion with the 90th percentile threshold

We create a dummy variable if gentrification is beyond the 75th percentile of this indicator:

```{r, warning=FALSE, error=FALSE}

# census %>%
# summarise(median_income_13=median(household_income_13, na.rm = T))

census <- census %>%
  mutate(one_gentri_4 = case_when(
    gentri_4 >= 0.095713333 & household_income_13 < 58676 ~ 1,
    household_income_13 > 58676 ~ 0,
    gentri_4 < 0.095713333 ~ 0
  ))
```

```{r, warning=FALSE, error=FALSE}
census %>%
  filter(one_gentri_4 == 1) %>%
  count()
```

This definition reaches 18 tracts.  


How many tracts coincide between definition 1 and 3?

```{r, warning=FALSE, error=FALSE}
census %>%
  mutate(common_gent = one_gentri_1 * one_gentri_3) %>%
  filter(common_gent == 1) %>%
  count()
```

```{r, warning=FALSE, error=FALSE}
census <- census %>%
  mutate(common_gent_1_3 = one_gentri_1 * one_gentri_3)
```

How many tracts coincide between definition 2 and 4?

```{r, warning=FALSE, error=FALSE}
census %>%
  mutate(common_gent = one_gentri_2 * one_gentri_4) %>%
  filter(common_gent == 1) %>%
  count()
```

```{r, warning=FALSE, error=FALSE}
census <- census %>%
  mutate(common_gent_2_4 = one_gentri_2 * one_gentri_4)
```

Save this new dataset:

```{r, warning=FALSE, error=FALSE}
saveRDS(census, file = "census.RDS")
```


#Summary Statistics in levels by gentrification definitions

```{r, warning=FALSE, error=FALSE}

census_income <- census %>%
  filter(household_income_13 < 58676.28)

census_group_13 <- census_income %>%
  select(
    household_income_13,
    income_pc_13,
    value_13,
    median_rent_13,
    share_white_only_13,
    share_black_not_latino_13,
    share_latino_13,
    population_13,
    one_gentri_1
  )

census_group_13 <- census_group_13[census_group_13$one_gentri_1 != "versicolor", ]

census_group_13$one_gentri_1 <- factor(census_group_13$one_gentri_1)

table_gent_1_level_13 <- tabular(Format(digits = 0) * household_income_13 + Format(digits = 0) * income_pc_13 + Format(digits = 0) * value_13 + Format(digits = 0) * median_rent_13 + Format(digits = 0) * population_13 + Format(digits = 2) * share_white_only_13 + Format(digits = 2) * share_black_not_latino_13 + Format(digits = 2) * share_latino_13 ~ one_gentri_1 * Justify(c, r) * ((n <- 1) + Median + Mean + std), data = census_group_13)

toLatex(table_gent_1_level_13)
table_gent_1_level_13
```

```{r, warning=FALSE, error=FALSE}
census_group_18 <- census_income %>%
  select(
    household_income_18,
    income_pc_18,
    value_18,
    median_rent_18,
    share_white_only_18,
    share_black_not_latino_18,
    share_latino_18,
    population_18,
    one_gentri_1
  )

census_group_18 <- census_group_18[census_group_18$one_gentri_1 != "versicolor", ]

census_group_18$one_gentri_1 <- factor(census_group_18$one_gentri_1)

table_gent_1_level_18 <- tabular(Format(digits = 0) * household_income_18 + Format(digits = 0) * income_pc_18 + Format(digits = 0) * value_18 + Format(digits = 0) * median_rent_18 + Format(digits = 0) * population_18 + Format(digits = 2) * share_white_only_18 + Format(digits = 2) * share_black_not_latino_18 + Format(digits = 2) * share_latino_18 ~ one_gentri_1 * Justify(c, r) * ((n <- 1) + Median + Mean + std), data = census_group_18)


# table_gent_1_level_18

toLatex(table_gent_1_level_18)
```


```{r, warning=FALSE, error=FALSE}

census_group_13 <- census_income %>%
  select(
    household_income_13,
    income_pc_13,
    value_13,
    median_rent_13,
    share_white_only_13,
    share_black_not_latino_13,
    share_latino_13,
    population_13,
    one_gentri_2
  )

census_group_13 <- census_group_13[census_group_13$one_gentri_2 != "versicolor", ]

census_group_13$one_gentri_2 <- factor(census_group_13$one_gentri_2)

table_gent_2_level_13 <- tabular(Format(digits = 0) * household_income_13 + Format(digits = 0) * income_pc_13 + Format(digits = 0) * value_13 + Format(digits = 0) * median_rent_13 + Format(digits = 0) * population_13 + Format(digits = 1) * share_white_only_13 + Format(digits = 1) * share_black_not_latino_13 + Format(digits = 1) * share_latino_13 ~ one_gentri_2 * Justify(c, r) * ((n <- 1) + Median + Mean + std), data = census_group_13)

toLatex(table_gent_2_level_13)
table_gent_2_level_13
```

```{r, warning=FALSE, error=FALSE}

census_group_18 <- census_income %>%
  select(
    household_income_18,
    income_pc_18,
    value_18,
    median_rent_18,
    share_white_only_18,
    share_black_not_latino_18,
    share_latino_18,
    population_18,
    one_gentri_2
  )

census_group_18 <- census_group_18[census_group_18$one_gentri_2 != "versicolor", ]

census_group_18$one_gentri_2 <- factor(census_group_18$one_gentri_2)

table_gent_2_level_18 <- tabular(Format(digits = 0) * household_income_18 + Format(digits = 0) * income_pc_18 + Format(digits = 0) * value_18 + Format(digits = 0) * median_rent_18 + Format(digits = 0) * population_18 + Format(digits = 1) * share_white_only_18 + Format(digits = 1) * share_black_not_latino_18 + Format(digits = 1) * share_latino_18 ~ one_gentri_2 * Justify(c, r) * ((n <- 1) + Median + Mean + std), data = census_group_18)

toLatex(table_gent_2_level_18)

table_gent_2_level_18
```


```{r, warning=FALSE, error=FALSE}

census_group_13 <- census_income %>%
  select(
    household_income_13,
    income_pc_13,
    value_13,
    median_rent_13,
    share_white_only_13,
    share_black_not_latino_13,
    share_latino_13,
    population_13,
    one_gentri_3
  )

census_group_13 <- census_group_13[census_group_13$one_gentri_3 != "versicolor", ]

census_group_13$one_gentri_3 <- factor(census_group_13$one_gentri_3)

table_gent_3_level_13 <- tabular(Format(digits = 0) * household_income_13 + Format(digits = 0) * income_pc_13 + Format(digits = 0) * value_13 + Format(digits = 0) * median_rent_13 + Format(digits = 0) * population_13 + Format(digits = 1) * share_white_only_13 + Format(digits = 1) * share_black_not_latino_13 + Format(digits = 1) * share_latino_13 ~ one_gentri_3 * Justify(c, r) * ((n <- 1) + Median + Mean + std), data = census_group_13)

toLatex(table_gent_3_level_13)
table_gent_3_level_13
```

```{r, warning=FALSE, error=FALSE}

census_group_18 <- census_income %>%
  select(
    household_income_18,
    income_pc_18,
    value_18,
    median_rent_18,
    share_white_only_18,
    share_black_not_latino_18,
    share_latino_18,
    population_18,
    one_gentri_3
  )

census_group_18 <- census_group_18[census_group_18$one_gentri_3 != "versicolor", ]

census_group_18$one_gentri_3 <- factor(census_group_18$one_gentri_3)

table_gent_3_level_18 <- tabular(Format(digits = 0) * household_income_18 + Format(digits = 0) * income_pc_18 + Format(digits = 0) * value_18 + Format(digits = 0) * median_rent_18 + Format(digits = 0) * population_18 + Format(digits = 1) * share_white_only_18 + Format(digits = 1) * share_black_not_latino_18 + Format(digits = 1) * share_latino_18 ~ one_gentri_3 * Justify(c, r) * ((n <- 1) + Median + Mean + std), data = census_group_18)

toLatex(table_gent_3_level_18)
table_gent_3_level_18
```



```{r, warning=FALSE, error=FALSE}

census_group_13 <- census_income %>%
  select(
    household_income_13,
    income_pc_13,
    value_13,
    median_rent_13,
    share_white_only_13,
    share_black_not_latino_13,
    share_latino_13,
    population_13,
    one_gentri_4
  )

census_group_13 <- census_group_13[census_group_13$one_gentri_4 != "versicolor", ]

census_group_13$one_gentri_4 <- factor(census_group_13$one_gentri_4)

table_gent_4_level_13 <- tabular(Format(digits = 0) * household_income_13 + Format(digits = 0) * income_pc_13 + Format(digits = 0) * value_13 + Format(digits = 0) * median_rent_13 + Format(digits = 0) * population_13 + Format(digits = 1) * share_white_only_13 + Format(digits = 1) * share_black_not_latino_13 + Format(digits = 1) * share_latino_13 ~ one_gentri_4 * Justify(c, r) * ((n <- 1) + Median + Mean + std), data = census_group_13)

toLatex(table_gent_4_level_13)
table_gent_4_level_13
```


```{r, warning=FALSE, error=FALSE}

census_group_18 <- census_income %>%
  select(
    household_income_18,
    income_pc_18,
    value_18,
    median_rent_18,
    share_white_only_18,
    share_black_not_latino_18,
    share_latino_18,
    population_18,
    one_gentri_4
  )

census_group_18 <- census_group_18[census_group_18$one_gentri_4 != "versicolor", ]

census_group_18$one_gentri_4 <- factor(census_group_18$one_gentri_4)

table_gent_4_level_18 <- tabular(Format(digits = 0) * household_income_18 + Format(digits = 0) * income_pc_18 + Format(digits = 0) * value_18 + Format(digits = 0) * median_rent_18 + Format(digits = 0) * population_18 + Format(digits = 1) * share_white_only_18 + Format(digits = 1) * share_black_not_latino_18 + Format(digits = 1) * share_latino_18 ~ one_gentri_4 * Justify(c, r) * ((n <- 1) + Median + Mean + std), data = census_group_18)

toLatex(table_gent_4_level_18)
table_gent_4_level_18
```


#Summary Statistics in changes by gentrification definitions


```{r, warning=FALSE, error=FALSE}

# census %>%
# summarise(median=median(household_income_13, na.rm=T))

census_group <- census_income %>%
  select(
    increase_value,
    increase_rent,
    increase_income,
    increase_income_pc,
    increase_share_latino,
    increase_share_black_not_latino,
    increase_pop,
    one_gentri_1
  )

census_group <- census_group[census_group$one_gentri_1 != "versicolor", ]

census_group$one_gentri_1 <- factor(census_group$one_gentri_1)

pvalue <- function(x) t.test(x ~ census_group$one_gentri_1)$p.value

table_gent_1 <- tabular(Format(digits = 2) * increase_value + Format(digits = 2) * increase_rent + Format(digits = 2) * increase_income + Format(digits = 2) * increase_income_pc + Format(digits = 2) * increase_pop + Format(digits = 2) * increase_share_latino + Format(digits = 2) * increase_share_black_not_latino ~ Format(digits = 2) * one_gentri_1 * Justify(c, r) * (Median + Mean + std) + pvalue, data = census_group)

toLatex(table_gent_1)
table_gent_1
```


```{r, warning=FALSE, error=FALSE}

# census %>%
# summarise(median=median(household_income_13, na.rm=T))

census_group <- census_income %>%
  select(
    increase_value,
    increase_rent,
    increase_income,
    increase_income_pc,
    increase_share_latino,
    increase_share_black_not_latino,
    increase_pop,
    one_gentri_2
  )

census_group <- census_group[census_group$one_gentri_2 != "versicolor", ]

census_group$one_gentri_2 <- factor(census_group$one_gentri_2)

pvalue <- function(x) t.test(x ~ census_group$one_gentri_2)$p.value

table_gent_2 <- tabular(Format(digits = 1) * increase_value + Format(digits = 1) * increase_rent + Format(digits = 1) * increase_income + Format(digits = 1) * increase_income_pc + Format(digits = 1) * increase_pop + Format(digits = 1) * increase_share_latino + Format(digits = 1) * increase_share_black_not_latino ~ Format(digits = 1) * one_gentri_2 * (Median + Mean + std) + pvalue, data = census_group)

toLatex(table_gent_2)
table_gent_2
```


Gentrification definition numer 2 (75th percentile):

```{r, warning=FALSE, error=FALSE}
census_group <- census_income %>%
  select(
    increase_value,
    increase_rent,
    increase_income,
    increase_income_pc,
    increase_share_latino,
    increase_share_black_not_latino,
    increase_pop,
    one_gentri_3
  )

census_group <- census_group[census_group$one_gentri_3 != "versicolor", ]

census_group$one_gentri_3 <- factor(census_group$one_gentri_3)

pvalue <- function(x) t.test(x ~ census_group$one_gentri_3)$p.value

table_gent_3 <- tabular(Format(digits = 1) * increase_value + Format(digits = 1) * increase_rent + Format(digits = 1) * increase_income + Format(digits = 1) * increase_income_pc + Format(digits = 1) * increase_pop + Format(digits = 1) * increase_share_latino + Format(digits = 1) * increase_share_black_not_latino ~ one_gentri_3 * Justify(c, r) * (Median + Mean + std) + pvalue, data = census_group)

toLatex(table_gent_3)

table_gent_3
```

Gentrification definition numer 2 (white 90th percentile):

```{r, warning=FALSE, error=FALSE}
census_group <- census_income %>%
  select(
    increase_value,
    increase_rent,
    increase_income,
    increase_income_pc,
    increase_share_latino,
    increase_share_black_not_latino,
    increase_pop,
    one_gentri_4
  )

census_group <- census_group[census_group$one_gentri_4 != "versicolor", ]

census_group$one_gentri_4 <- factor(census_group$one_gentri_4)

pvalue <- function(x) t.test(x ~ census_group$one_gentri_4)$p.value

table_gent_4 <- tabular(Format(digits = 1) * increase_value + Format(digits = 1) * increase_rent + Format(digits = 1) * increase_income + Format(digits = 1) * increase_income_pc + Format(digits = 1) * increase_pop + Format(digits = 1) * increase_share_latino + Format(digits = 1) * increase_share_black_not_latino ~ Format(digits = 1) * one_gentri_4 * Justify(c, r) * (Median + Mean + std) + pvalue, data = census_group)

toLatex(table_gent_4)

table_gent_4
```


There is a positive correlation between the continuous definition of gentrification 1 and 2. 

```{r, warning=FALSE, error=FALSE}
ggplot(data = census) +
  geom_point(mapping = aes(x = gentri_3, y = gentri_4)) +
  geom_smooth(mapping = aes(x = gentri_3, y = gentri_4))
```



```{r, warning=FALSE, error=FALSE}
gent <- census
```

```{r, warning=FALSE, error=FALSE}
saveRDS(gent, file = "gent.RDS")
```

#Current legislation

How many Tracts qualify under current legistlation (assumming that the increases compound)?

```{r, warning=FALSE, error=FALSE}
census <- census %>%
  mutate(increase_value_annual=(1+increase_value)^(1/5)-1)

census %>%
  summarise(sum(population_18))


census_legis <- census %>%
  filter(increase_value_annual>0.10 & (household_income_18>75000 &   household_income_18<100000) | 
           increase_value_annual>0.07 & household_income_18<75000)

census_legis %>%
  count()
  
census_legis %>%
  summarise(sum(population_18))

```

25 out of 716 tracts qualify under current legislation (almost 3.5%) assuming that all have 10 years of residence. The total population of these tracts is 70,464 out of 2,632,842 that equals to a 2.6% in our sample.  


How many Tracts qualify after the changes we want in legistlation (assumming that the increases compound)?

I will use AMI for April 2019 (https://www.chicago.gov/city/en/depts/dcd/supp_info/area_median_incomeamichart.html).

```{r, warning=FALSE, error=FALSE}

census_legis <- census %>%
  filter(increase_value_annual>0.5 & (household_income_18>64000 & household_income_18<80000) | 
           increase_value_annual>0.03 & household_income_18<64000)

census_legis %>%
  count()
  
census_legis %>%
  summarise(sum(population_18))

```


How many Tracts qualify under current legistlation (assumming that the increases do not compound)?


```{r, warning=FALSE, error=FALSE}
census <- census %>%
  mutate(increase_value_not_compund=increase_value/5)


census_legis <- census %>%
  filter(increase_value_not_compund>0.10 & (household_income_18>75000 &   household_income_18<100000) | 
           increase_value_not_compund>0.07 & household_income_18<75000)

census_legis %>%
  count()
  
census_legis %>%
  summarise(sum(population_18))

```

29 out of 716 tracts qualify under current legislation (4%) assuming that all have 10 years of residence. The total population of these tracts is 82,988 out of 2,632,842 that equals to a 3.15% in our sample.  

How many Tracts qualify after the changes we want in legistlation (assumming that the increases compound)?

I will use AMI for April 2019 (https://www.chicago.gov/city/en/depts/dcd/supp_info/area_median_incomeamichart.html).

```{r, warning=FALSE, error=FALSE}

census_legis <- census %>%
  filter(increase_value_not_compund>0.5 & (household_income_18>64000 & household_income_18<80000) | 
           increase_value_not_compund>0.03 & household_income_18<64000)

census_legis %>%
  count()
  
census_legis %>%
  summarise(sum(population_18))

```

