---
title: "Statistics on Gentrification"
author: "Fernando Garcia - Drazzel Feliu"
date: "8/25/2020"
output: html_document
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(lubridate)
library(ggmap)
library(tidycensus)
library(mapview)
library(sf)
library(leaflet)
library(viridis)
library(maps)
library(htmlwidgets)
library(htmltools)
library(tigris)
library(reshape2)
library(ggrepel)
library(ggmap)
library(tables)

rm(list = ls())
```

```{r, warning=FALSE, error=FALSE}

census <- readRDS("census.RDS")
```

Checking summary statistics:

```{r, warning=FALSE, error=FALSE}
census %>%
  mutate(income_pc_18 = agg_income_18 / population_18) %>%
  summarise(mean_income_pc_18 = mean(income_pc_18, na.rm = T))

census %>%
  summarise(mean_avg_household_size_18 = mean(avg_household_size_18, na.rm = T))

census %>%
  mutate(income_household_18 = (agg_income_18 / population_18) *
    avg_household_size_18) %>%
  summarise(median_income_household_18 = median(income_household_18, na.rm = T))

census %>%
  summarise(median_income_18 = median(income_18, na.rm = T))

census %>%
  summarise(median_rent_18 = median(median_rent_18, na.rm = T))
```

Mean income per capita for 2018 is $33,599. Median household income is \$68,000 with an average household size of 2.6. Correct if we look at https://www.census.gov/quickfacts/fact/table/chicagocityillinois/LND110210.

Summary statistics for 2013:

```{r, warning=FALSE, error=FALSE}

census %>%
  mutate(income_pc_13 = agg_income_13 / population_13) %>%
  summarise(mean_income_pc_13 = mean(income_pc_13, na.rm = T))

census %>%
  summarise(mean_avg_household_size_13 = mean(avg_household_size_13, na.rm = T))

census %>%
  mutate(income_household_13 = (agg_income_13 / population_13) *
    avg_household_size_13) %>%
  summarise(median_income_household_13 = median(income_household_13, na.rm = T))
```

Median household income is \$ 58,676.28 and per capita income mean is \$28,747.67 in 2013.


#Gentrification:

Distribution of home values in 2018:

```{r}

census %>%
  summarise(median_value_18 = median(value_18, na.rm = T))

ggplot(census) +
  geom_histogram(aes(x = value_18, y = ..density..),
    binwidth = 10000,
    colour = "black", fill = "white"
  ) +
  geom_vline(xintercept = 236600, color = "blue") +
  ggtitle("Histogram price increase in home values")
``` 

Distribution of home value increase:

```{r}

# census %>%
# summarise(median_value_increase=median(increase_value, na.rm=T))

ggplot(census) +
  geom_histogram(aes(x = increase_value, y = ..density..),
    binwidth = 1,
    colour = "black", fill = "white"
  ) +
  geom_vline(xintercept = 4.437273, color = "blue") +
  ggtitle("Histogram price increase home values")
``` 

We will zoom in to look at the outliers:

```{r}
ggplot(census) +
  geom_histogram(aes(x = increase_value, y = ..density..),
    binwidth = 1,
    colour = "black", fill = "white"
  ) +
  xlim(c(25, 200)) +
  ggtitle("Histogram price increase in home values")
``` 

```{r, warning=FALSE, error=FALSE}
census_rest <- census %>%
  filter(increase_value > 0)

plot(ecdf(census_rest$increase_value),
  xlab = "Home value price increase ($)",
  ylab = "Cumulative Distribution",
  main = "Home value price increase distributions from 2013-2018"
)
abline(h = 0.95, col = "blue")
```


The cumulative distribution looks like this:

```{r, warning=FALSE, error=FALSE}
homeincrease <- census$increase_value

quantile(homeincrease, c(.5, .75, .8, .85, .9, .95), na.rm = T)
```

The last 10% of the distribution gets a 36% or more of an increase in home values. We can use this to define gentrification.

#First definition of gentrification

We use the 90th percentile of the home value increase 

Significant increases in home values should be reflective of some broader sense of gentrification. As the neighborhood gentrifies, the value of property within the area should be increasing as wealthier individuals begin to displace the individuals living in the area prior. Using the defined distribution of changes in home values between 2013 and 2018, we use the 80% and 90% threshholds to identify gentrification patterns across the city.

```{r, warning=FALSE, error=FALSE}

census <- census %>%
  mutate(gentri_1 = case_when(
    increase_value > 36 & value_18 > 180800 ~ 1,
    increase_value > 25 & value_18 > 226000 ~ 1,
    increase_value <= 36 & (value_18 > 180800 & value_18 < 226000) ~ 0,
    (increase_value > 25 & increase_value <= 36) & value_18 <= 226000 ~ 0,
    increase_value <= 25 ~ 0,
    value_18 <= 180800 ~ 0
  ))
```

```{r, warning=FALSE, error=FALSE}

census %>%
  filter(gentri_1 == 1) %>%
  count()
```

This measure of gentrification reaches 76 tracts.


#Second definition of gentrification:

Gentrification isn't simply the increase in wealth for a neighborhood. We can confirm this by checking the demographic makeup of neighborhoods who have seen median income increases. This definition of gentrification then serves to identify the neighborhoods where were median incomes have changed drastically in a 5 year period.

```{r, warning=FALSE, error=FALSE}

census <- census %>%
  mutate(gentri_2 = case_when(
    increase_value > 36 & household_income_13 <= 58676 ~ 1,
    increase_value > 25 & household_income_13 <= 46900 ~ 1,
    (increase_value > 25 & increase_value <= 36) &
      (household_income_13 > 46900 & household_income_13 < 58676) ~ 0,
    increase_value <= 25 ~ 0,
    household_income_13 > 58676 ~ 0
  ))

# I am not convinced of the value of this definition using both home value increase and level of income.
```

```{r, warning=FALSE, error=FALSE}

census %>%
  filter(gentri_2 == 1) %>%
  count()
```

This measure of gentrification reaches 35 tracts.

#Gentrification definition according to Brummet, Reed, 2019. 

```{r, warning=FALSE, error=FALSE}
census <- census %>%
  mutate(gentri_3 = (bc_18 - bc_13) / population_13)
```

```{r}

census %>%
  summarise(median_gentri_3 = median(gentri_3, na.rm = T))

ggplot(census) +
  geom_histogram(aes(x = gentri_3, y = ..density..),
    colour = "black", fill = "white"
  ) +
  geom_vline(xintercept = 0.02742509, color = "blue") +
  ggtitle("Histogram price increase in home values")
``` 

What is the 90th percentile of the gentrification's indicator:

```{r, warning=FALSE, error=FALSE}
gentrification <- census$gentri_3

quantile(gentrification, c(.5, .75, .8, .85, .90, .95), na.rm = T)
```

We create a dummy variable if gentrification is beyond the 90th percentile of this indicator:

```{r, warning=FALSE, error=FALSE}

# census %>%
# summarise(median_income_13=median(household_income_13, na.rm = T))

census <- census %>%
  mutate(one_gentri_3 = case_when(
    gentri_3 >= 0.06704281 & household_income_13 < 58676 ~ 1,
    household_income_13 > 58676 ~ 0,
    gentri_3 < 0.06704281 ~ 0
  ))
```

```{r, warning=FALSE, error=FALSE}
census %>%
  filter(one_gentri_3 == 1) %>%
  count()
```

This measure of gentrification reaches 22 tracts.



Gentrification definition number 4 using racial data:

```{r, warning=FALSE, error=FALSE}

census <- census %>%
  mutate(gentri_4 = (white_not_latino_18 - white_not_latino_13) / sum_13)
```


```{r}

# census %>%
# summarise(median_gentri_4=median(gentri_4, na.rm=T))

ggplot(census) +
  geom_histogram(aes(x = gentri_4, y = ..density..),
    colour = "black", fill = "white"
  ) +
  geom_vline(xintercept = 0.002240093, color = "blue") +
  ggtitle("Histogram price increase in home values")
``` 

What is the 90th percentile of the gentrification's indicator:

```{r, warning=FALSE, error=FALSE}
gentrification <- census$gentri_4

quantile(gentrification, c(.5, .75, .8, .85, .90, .95), na.rm = T)
```

We create a dummy variable if gentrification is beyond the 90th percentile of this indicator:

```{r, warning=FALSE, error=FALSE}

# census %>%
# summarise(median_income_13=median(household_income_13, na.rm = T))

census <- census %>%
  mutate(one_gentri_4 = case_when(
    gentri_4 >= 0.038203177 & household_income_13 < 58676 ~ 1,
    household_income_13 > 58676 ~ 0,
    gentri_4 < 0.038203177 ~ 0
  ))
```

```{r, warning=FALSE, error=FALSE}
census %>%
  filter(one_gentri_4 == 1) %>%
  count()
```

Definition number 4 reaches 18 tracts.




How many tracts coincide between definition 3 and 4?

```{r, warning=FALSE, error=FALSE}
census %>%
  mutate(common_gent = one_gentri_3 * one_gentri_4) %>%
  filter(common_gent == 1) %>%
  count()
```

```{r, warning=FALSE, error=FALSE}
census <- census %>%
  mutate(common_gent_3_4 = one_gentri_3 * one_gentri_4)
```

```{r, warning=FALSE, error=FALSE}
saveRDS(census, file = "census.RDS")
```


Summary statistics for gentrified areas according to definition 3:


```{r, warning=FALSE, error=FALSE}

census_income <- census %>%
  filter(household_income_13 < 58676)

aggregate(cbind(
  y = census_income$household_income_18,
  rent = census_income$median_rent_18,
  black = census_income$share_black_18,
  black_13 = census_income$share_black_13,
  latino = census_income$share_latino_18,
  latino_13 = census_income$share_latino_13,
  change_p = census_income$increase_value,
  change_y = census_income$increase_income,
  change_rent = census_income$increase_rent,
  rent_y_18 = census_income$median_rent_income_18,
  rent_y_13 = census_income$median_rent_income_13
),
# Specify data column
by = list(census_income$one_gentri_3), # Specify group indicator
FUN = mean, na.rm = T
)

t.test(increase_value ~ one_gentri_3, data = census_income)
t.test(increase_rent ~ one_gentri_3, data = census_income)
```

Summary statistics for gentrified areas according to definition 4:


```{r, warning=FALSE, error=FALSE}

census_income <- census %>%
  filter(household_income_13 < 58676)

aggregate(cbind(
  y = census_income$household_income_18,
  rent = census_income$median_rent_18,
  black_13 = census_income$share_black_13,
  black = census_income$share_black_18,
  latino_13 = census_income$share_latino_13,
  latino = census_income$share_latino_18,
  change_p = census_income$increase_value,
  change_y = census_income$increase_income,
  change_rent = census_income$increase_rent,
  rent_y_18 = census_income$median_rent_income_18,
  rent_y_13 = census_income$median_rent_income_13
), # Specify data column
  by = list(census_income$one_gentri_4), # Specify group indicator
  FUN = mean, 
  na.rm = T
)

t.test(increase_value ~ one_gentri_4, data = census_income)
t.test(increase_rent ~ one_gentri_4, data = census_income)

```

```{r, warning=FALSE, error=FALSE}

census_group <- census_income %>%
  select(
    increase_value, increase_rent, increase_income, increase_latino, increase_black, share_latino_13, share_black_13,
    share_latino_18, share_black_18, one_gentri_3
  )

census_group <- census_group[census_group$one_gentri_3 != "versicolor", ]

census_group$one_gentri_3 <- factor(census_group$one_gentri_3)

census_group <- na.omit(census_group, cols = "increase_value")


P.value <- function(x) t.test(x ~ census_group$one_gentri_3)$p.value

table_gent_3 <- tabular(increase_value + increase_rent + increase_income +increase_latino + increase_black + share_latino_13
  + share_latino_18 + share_black_13  + share_black_18 ~ one_gentri_3 * (mean + sd) +P.value, data = census_group)

table_gent_3

```

```{r, warning=FALSE, error=FALSE}

census_group <- census_income %>%
  select(
    increase_value, increase_rent, increase_income, share_latino_13, share_black_13,
    share_latino_18, share_black_18, one_gentri_4
  )

census_group <- census_group[census_group$one_gentri_4 != "versicolor", ]

census_group$one_gentri_4 <- factor(census_group$one_gentri_4)

census_group <- na.omit(census_group, cols = "increase_value")


P.value <- function(x) t.test(x ~ census_group$one_gentri_4)$p.value

table_gent_4 <- tabular(increase_value + increase_rent + increase_income + share_latino_13
  + share_latino_18 + share_black_13  + share_black_18 ~ one_gentri_4 * (mean + sd) +P.value, data = census_group)

table_gent_4

```



```{r, warning=FALSE, error=FALSE}
gent <- census
```

```{r, warning=FALSE, error=FALSE}
saveRDS(gent, file = "gent.RDS")
```
