---
title: "Data Visualization"
author: "Fernando Garcia - Drazzel Feliu"
date: "8/25/2020"
output: html_document
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(lubridate)
library(ggmap)
library(tidycensus)
library(mapview)
library(sf)
library(leaflet)
library(viridis)
library(maps)
library(htmlwidgets)
library(htmltools)
library(tigris)
library(reshape2)
library(ggrepel)
library(ggmap)
library(snakecase)
library(scales)

rm(list = ls())
```

```{r, warning=FALSE, error=FALSE}

gent <- readRDS("gent.RDS")



```

Neighborhood Polygons:

```{r, warning=FALSE, error=FALSE}
community <- read_sf("geo_export_72cf3d72-fa82-4448-a382-c9dd6dcb4fd2.shp")

community <- st_transform(
  community,
  4326
)

community$centroid <- st_centroid(community$geometry)

coords <- do.call(rbind, st_geometry(community$centroid)) %>% 
    as_tibble() %>% setNames(c("lon","lat"))

community$lon <- coords$lon
community$lat <- coords$lat
```

Leaflet Options and Colors:

```{r Leaflet Options - Community Areas}
neigh_h_lights <- highlightOptions(
  weight = 4.5,
  color = "#ffffff",
  fillOpacity = .5,
  bringToFront = TRUE
)

neigh_l_options <- labelOptions(
  style = list("font-weight" = "normal", padding = "3px 8px"),
  textsize = "15px",
  direction = "auto"
)

neigh_labels <- sprintf(
  paste0(
    "<strong>Name: </strong>%s<br/>"
  ),
  to_title_case(community$community)
) %>%
  lapply(HTML)
```

```{r Leaflet Options - Census Tracts}
tract_h_lights <- highlightOptions(
  weight = 4.5,
  color = "#ffffff",
  fillOpacity = .5,
  bringToFront = TRUE
)

tract_l_options <- labelOptions(
  style = list("font-weight" = "normal", padding = "3px 8px"),
  textsize = "15px",
  direction = "auto"
)

gent$tract_labels <- sprintf(
  paste0(
    "<strong>Name: </strong>%s<br/>",
    "<strong>2013 Data:</strong><br/>
    <strong>Population: </strong>%g<br/>",
    "<strong>Median Household Income: </strong>%s<br/>",
    "<strong>Median Housing Value: </strong>%s<br/>",
    "<strong>Percentage of Population With A Bachelor's Degree or Higher: </strong>%g%%<br/>",
    "<strong>Racial Demographics</strong><br/><strong>White: </strong>%g%%<br/><strong>Black: </strong>%g%%<br/><strong>Latino: </strong>%g%%<br/>",
    "<br/> <strong>2018 Data:</strong><br/>
    <strong>Population: </strong>%g<br/>",
    "<strong>Median Household Income: </strong>%s<br/>",
    "<strong>Median Housing Value: </strong>%s<br/>",
    "<strong>Percentage of Population With A Bachelor's Degree or Higher: </strong>%g%%<br/>",
    "<strong>Racial Demographics</strong><br/><strong>White: </strong>%g%%<br/><strong>Black: </strong>%g%%<br/><strong>Latino: </strong>%g%%<br/>",
    "<br> <strong>Changes:</strong><br/><strong>Median Household Income: </strong>%g%%"
  ),
  gent$NAME,
  gent$population_13,
  dollar(gent$income_13),
  dollar(gent$value_13),
  round((gent$bc_13/gent$population_13)*100,2),
  round((gent$share_white_only_13)*100,2),
  round((gent$share_black_not_latino_13)*100,2),
  round((gent$share_latino_13)*100,2),
  gent$population_18,
  dollar(gent$income_18),
  dollar(gent$value_18),
  round((gent$bc_18/gent$population_18)*100,2),
  round((gent$share_white_only_18)*100,2),
  round((gent$share_black_not_latino_18)*100,2),
  round((gent$share_latino_18)*100,2),
  round((gent$increase_income)*100,2)
) %>%
  lapply(HTML)
```

#Different Maps with increase in property values as example

Using ggplot:

```{r, warning=FALSE, error=FALSE}
ggplot() +
  geom_sf(gent,
    mapping = aes(
      fill = increase_value,
      color = increase_value
    ),
    inherit.aes = F,
    alpha = .99
  ) +
  scale_fill_viridis() +
  scale_color_viridis() +
  theme_minimal() +
  ggtitle("Percentage increase in property values between 2018 and 2013", )
```

```{r, warning=FALSE, error=FALSE}
ggplot() +
  geom_sf(gent,
    mapping = aes(
      fill = share_black_13,
      color = share_black_13
    ),
    inherit.aes = F,
    alpha = .99
  ) +
  scale_fill_viridis() +
  scale_color_viridis() +
  theme_minimal() +
  ggtitle("Share of black population by tracts in 2013", )
```

```{r, warning=FALSE, error=FALSE}
ggplot() +
  geom_sf(gent,
    mapping = aes(
      fill = share_black_18,
      color = share_black_18
    ),
    inherit.aes = F,
    alpha = .99
  ) +
  scale_fill_viridis() +
  scale_color_viridis() +
  theme_minimal() +
  ggtitle("Share of black population by tracts in 2018", )
```

```{r, warning=FALSE, error=FALSE}
ggplot() +
  geom_sf(gent,
    mapping = aes(
      fill = share_latino_18,
      color = share_latino_18
    ),
    inherit.aes = F,
    alpha = .99
  ) +
  scale_fill_viridis() +
  scale_color_viridis() +
  theme_minimal() +
  ggtitle("Share of latino population by tracts in 2018", )
```

Visualizations begin

```{r, warning=FALSE, error=FALSE}

color_200k <- colorNumeric(
  palette = c("red","white","green"),
  domain = c(-max(gent$increase_value,na.rm = T),max(gent$increase_value,na.rm = T)))

gent$color_inc_value <- color_200k(gent$increase_value)
```

Map #3:


```{r}
leaflet(options = leafletOptions(zoomSnap = 0.1)) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    data = st_geometry(community),
    color = "black",
    fillColor = "gray",
    popup = neigh_labels,
    popupOptions = neigh_l_options,
    highlightOptions = neigh_h_lights,
    fillOpacity = .1,
    weight = 3,
    group = "Community Areas"
  ) %>% 
  addPolygons(
    data = st_geometry(gent),
    color = "black",
    fillColor = gent$color_inc_value,
    popup = gent$tract_labels,
    popupOptions = tract_l_options,
    highlightOptions = tract_h_lights,
    weight = 1,
    fillOpacity = .5,
    smoothFactor = .2,
    group = "Tracts"
  ) %>%
  addLegend(
    title = "Percentage change housing values<br> between 2018 and 2013",
    pal = color_200k,
    values = gent$increase_value,
    position = "topright"
  ) %>% 
  addLayersControl(
    overlayGroups = c("Tracts", "Community Areas"),
    position = "bottomright"
  )
```

Change in percentage of population with a bachelor's degree is likely correlated to increases in median household income. Is change in degree also correlated with the growth in housing value?

What makes up the increase income variable?

Add CTA map overlay to viz.

Continuous third definition of gentrification: 

```{r, warning=FALSE, error=FALSE}

color_200k <- colorNumeric(
  palette = c("white", "blue"),
  domain = gent$gentri_3
)
```

```{r}

leaflet(options = leafletOptions(zoomSnap = 0.1)) %>%
  addTiles() %>%
  addPolygons(
    data = st_geometry(gent),
    color = "black",
    fillColor = color_200k(gent$gentri_3),
    weight = 2.5,
    fillOpacity = .3,
    smoothFactor = .2,
    group = "Gentrification By Education"
  ) %>%
  addPolygons(
    data = st_geometry(community),
    popup = neigh_labels,
    popupOptions = neigh_l_options,
    highlightOptions = neigh_h_lights,
    group = "Community Areas"
  ) %>%
  addLegend(
    title = "Gentrified tracts between 2018 and 2013",
    pal = color_200k,
    values = gent$gentri_3,
    position = "topright"
  ) %>%
  addLayersControl(
    overlayGroups = c("Gentrification By Education", "Community Areas"),
    position = "bottomright"
  )
```



These are the tracts that would be gentrified according to definition 3:

```{r}

leaflet(options = leafletOptions(zoomSnap = 0.1)) %>%
  addTiles() %>%
  addPolygons(
    data = st_geometry(gent %>% filter(gent$one_gentri_1 == 1)),
    color = "black",
    fillColor = color_200k(gent$gentri_3),
    weight = 2.5,
    fillOpacity = .3,
    smoothFactor = .2,
    popup = gent$tract_labels,
    popupOptions = tract_l_options,
    highlightOptions = tract_h_lights,
    group = "Gentrification By Education"
  )  %>%
  addPolygons(
    data = st_geometry(community),
    popup = neigh_labels,
    popupOptions = neigh_l_options,
    highlightOptions = neigh_h_lights,
    group = "Community Areas"
  ) %>%
  addLayersControl(
    overlayGroups = c("Gentrification By Education", "Community Areas"),
    position = "bottomright"
  )
```

These are the tracts that are gentrified according to definition 4:

```{r}

leaflet(options = leafletOptions(zoomSnap = 0.1)) %>%
  addTiles() %>%
  addPolygons(
    data = st_geometry(gent %>% filter(gent$one_gentri_3 == 1)),
    color = "black",
    fillColor = color_200k(gent$increase_income),
    weight = 2.5,
    fillOpacity = .3,
    smoothFactor = .2,
    popup = tract_labels,
    popupOptions = tract_l_options,
    highlightOptions = tract_h_lights,
    group = "Gentrification By Education"
  ) %>%
  addPolygons(
    data = st_geometry(community),
    popup = neigh_labels,
    popupOptions = neigh_l_options,
    highlightOptions = neigh_h_lights,
    group = "Community Areas"
  ) %>%
  addLayersControl(
    overlayGroups = c("Gentrification By Education", "Community Areas"),
    position = "bottomright"
  )
```

Gentrified tracts according to definitions 1 (blue) and 3 (black) with tracts that coincide in red: 

```{r}

leaflet(options = leafletOptions(zoomSnap = 0.1)) %>%
  addTiles() %>%
  addPolygons(
    data = st_geometry(gent %>% filter(gent$one_gentri_1 == 1)),
    color = "blue",
    fillColor = "blue",
    weight = 2.5,
    fillOpacity = .3,
    smoothFactor = .2,
    popup = tract_labels,
    popupOptions = tract_l_options,
    highlightOptions = tract_h_lights,
    group = "Gentrification By Education"
  ) %>%
  addPolygons(
    data = st_geometry(gent %>% filter(gent$one_gentri_3 == 1)),
    color = "black",
    fillColor = color_200k(gent$increase_income),
    weight = 2.5,
    fillOpacity = .3,
    smoothFactor = .2,
    popup = tract_labels,
    popupOptions = tract_l_options,
    highlightOptions = tract_h_lights,
    group = "Gentrification By Education"
  ) %>%
  addPolygons(
    data = st_geometry(gent %>% filter(gent$common_gent_1_3 == 1)),
    color = "black",
    fillColor = "red",
    weight = 2.5,
    fillOpacity = .3,
    smoothFactor = .2,
    popup = tract_labels,
    popupOptions = tract_l_options,
    highlightOptions = tract_h_lights,
    group = "Gentrification By Education"
  )  %>%
  addPolygons(
    data = st_geometry(community),
    popup = neigh_labels,
    popupOptions = neigh_l_options,
    highlightOptions = neigh_h_lights,
    group = "Community Areas"
  ) %>%
  addLayersControl(
    overlayGroups = c("Gentrification By Education", "Community Areas"),
    position = "bottomright"
  )
```

```{r}

leaflet(options = leafletOptions(zoomSnap = 0.1)) %>%
  addTiles() %>%
  addPolygons(
    data = st_geometry(gent %>% filter(gent$one_gentri_1 == 1 & increase_value >15)),
    color = "black",
    fillColor = "red",
    weight = 2.5,
    fillOpacity = .3,
    smoothFactor = .2,
    popup = tract_labels,
    popupOptions = tract_l_options,
    highlightOptions = tract_h_lights,
    group = "Gentrification By Education"
  )  %>%
  addPolygons(
    data = st_geometry(gent %>% filter(gent$one_gentri_3 == 1 & increase_value >15)),
    color = "black",
    fillColor = "blue",
    weight = 2.5,
    fillOpacity = .3,
    smoothFactor = .2,
    popup = tract_labels,
    popupOptions = tract_l_options,
    highlightOptions = tract_h_lights,
    group = "Gentrification By Education"
  ) %>%
  addPolygons(
    data = st_geometry(community),
    popup = neigh_labels,
    popupOptions = neigh_l_options,
    highlightOptions = neigh_h_lights,
    group = "Community Areas"
  ) %>%
  addLayersControl(
    overlayGroups = c("Gentrification By Education", "Community Areas"),
    position = "bottomright"
  )
```

--------------------------------
OUTPUT MAPS
--------------------------------

```{r Neighborhoods}
leaflet(options = leafletOptions(zoomSnap = 0.1)) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    data = st_geometry(community),
    color = "black",
    fillColor = "gray",
    popup = neigh_labels,
    fillOpacity = .1,
    weight = 3,
    group = "Community Areas"
  ) %>% 
  addLabelOnlyMarkers(
    lng = community$lon, lat = community$lat,
    label = to_title_case(substr(community$community, 1, 6)),
    labelOptions = labelOptions(noHide = T, textOnly = T, textsize = "6px",
                                direction = "center",
                                style = list(
                                  "color" = "red"
                                )))
```

```{r Neighborhoods & Tracts}
leaflet(options = leafletOptions(zoomSnap = 0.1)) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    data = st_geometry(community),
    color = "black",
    fillColor = "gray",
    popup = neigh_labels,
    fillOpacity = .1,
    weight = 3,
    group = "Community Areas"
  ) %>% 
  addPolygons(
    data = st_geometry(gent),
    color = "black",
    fillColor = "gray",
    fillOpacity = .1,
    weight = 1
  )
```

```{r Median Income Level 2013}

median_income <- colorNumeric(
  palette = "magma",
  domain = gent$income_13)

leaflet(options = leafletOptions(zoomSnap = 0.1)) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    data = st_geometry(community),
    color = "black",
    fillColor = "gray",
    popup = neigh_labels,
    fillOpacity = .1,
    weight = 5,
    group = "Community Areas"
  ) %>% 
  addPolygons(
    data = st_geometry(gent),
    color = "black",
    fillColor = median_income(gent$income_13),
    fillOpacity = .6,
    weight = 1
  ) %>% 
  addLegend(
    title = "Median Incomes in 2013",
    pal = median_income,
    values = gent$income_13,
    position = "topright"
  ) 
```

```{r Median Home Value in 2013}
median_home_value <- colorNumeric(
  palette = "viridis",
  domain = gent$value_13)

leaflet(options = leafletOptions(zoomSnap = 0.1)) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    data = st_geometry(community),
    color = "black",
    fillColor = "gray",
    popup = neigh_labels,
    fillOpacity = .1,
    weight = 5,
    group = "Community Areas"
  ) %>% 
  addPolygons(
    data = st_geometry(gent),
    color = "black",
    fillColor = median_home_value(gent$value_13),
    fillOpacity = .6,
    weight = 1
  ) %>% 
  addLegend(
    title = "Median Home Value in 2013",
    pal = median_home_value,
    values = gent$value_13,
    position = "topright"
  ) 
```

```{r Tracts Under Gentrification Definition 1}
leaflet(options = leafletOptions(zoomSnap = 0.1)) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    data = st_geometry(community),
    color = "black",
    fillColor = "black",
    popup = neigh_labels,
    fillOpacity = .1,
    weight = 5,
    group = "Community Areas"
  ) %>% 
  addPolygons(
    data = st_geometry(gent),
    color = "black",
    fillColor = ifelse(gent$one_gentri_2==1,"red",
                       ifelse(gent$one_gentri_1==1,"blue","white")),
    fillOpacity = .6,
    weight = 1
  ) %>% 
  addLegend(
    title = "Gentrification defitintion 1",
    colors =c("blue",  "red"),
    labels= c("75th percentile", "75th and 90th percentile"),
    position = "topright"
  ) 
```

```{r Tracts Under Gentrification Definition 2}
leaflet(options = leafletOptions(zoomSnap = 0.1)) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    data = st_geometry(community),
    color = "black",
    fillColor = "gray",
    popup = neigh_labels,
    fillOpacity = .1,
    weight = 5,
    group = "Community Areas"
  ) %>% 
  addPolygons(
    data = st_geometry(gent),
    color = "black",
    fillColor = ifelse(gent$one_gentri_4==1,"red",
                       ifelse(gent$one_gentri_3==1,"blue","white")),
    fillOpacity = .6,
    weight = 1
  ) %>% 
  addLegend(
    title = "Gentrification defitintion 2",
    colors =c("blue",  "red"),
    labels= c("75th percentile", "75th and 90th percentile"),
    position = "topright"
  ) 
```

```{r Tracts Under Both Gentrification Definitions at 75th percentile}
leaflet(options = leafletOptions(zoomSnap = 0.1)) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    data = st_geometry(community),
    color = "black",
    fillColor = "black",
    popup = neigh_labels,
    fillOpacity = .1,
    weight = 5,
    group = "Community Areas"
  ) %>% 
  addPolygons(
    data = st_geometry(gent),
    color = "black",
    fillColor = ifelse(gent$one_gentri_1*gent$one_gentri_3==1,"red","white"),
    fillOpacity = .6,
    weight = 1
  )
```


```{r Increase Median Income by defintion 1 - Slide 22}
census_income <- gent %>% filter(one_gentri_1!=0)

median_increase_income <- colorNumeric(
  palette = "viridis",
  domain = census_income$increase_income)

leaflet(options = leafletOptions(zoomSnap = 0.1)) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    data = st_geometry(community),
    color = "black",
    fillColor = "gray",
    popup = neigh_labels,
    fillOpacity = .1,
    weight = 5,
    group = "Community Areas"
  ) %>% 
  addPolygons(
    data = st_geometry(census_income),
    color = "black",
    fillColor = median_increase_income(census_income$increase_income),
    fillOpacity = .6,
    weight = 1
  ) %>% 
  addLegend(
    title = "Increase in Median Household Income",
    pal = median_increase_income,
    values = census_income$increase_income,
    position = "topright"
  ) 
```

```{r Increase Median House Value by defintion 1 - Slide 23}
median_increase_value <- colorNumeric(
  palette = "viridis",
  domain = census_income$increase_value)

leaflet(options = leafletOptions(zoomSnap = 0.1)) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    data = st_geometry(community),
    color = "black",
    fillColor = "gray",
    popup = neigh_labels,
    fillOpacity = .1,
    weight = 5,
    group = "Community Areas"
  ) %>% 
  addPolygons(
    data = st_geometry(census_income),
    color = "black",
    fillColor = median_increase_value(census_income$increase_value),
    fillOpacity = .6,
    weight = 1
  ) %>% 
  addLegend(
    title = "Increase in Home Value",
    pal = median_increase_value,
    values = census_income$increase_value,
    position = "topright"
  ) 

```

PLEASE NOTE: This following section must be run sequentially after the previous section. Running chunks out of order will result in wrong maps.

```{r Increase Median Income by defintion 2 Slide 25}
census_income <- gent %>% filter(one_gentri_3!=0)

median_increase_income <- colorNumeric(
  palette = "viridis",
  domain = census_income$increase_income)

leaflet(options = leafletOptions(zoomSnap = 0.1)) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    data = st_geometry(community),
    color = "black",
    fillColor = "gray",
    popup = neigh_labels,
    fillOpacity = .1,
    weight = 5,
    group = "Community Areas"
  ) %>% 
  addPolygons(
    data = st_geometry(census_income),
    color = "black",
    fillColor = median_increase_income(census_income$increase_income),
    fillOpacity = .6,
    weight = 1
  ) %>% 
  addLegend(
    title = "Increase in Household Income",
    pal = median_increase_income,
    values = census_income$increase_income,
    position = "topright"
  ) 
```

```{r Increase Median House Value by defintion 2 Slide 26}
median_increase_value <- colorNumeric(
  palette = "viridis",
  domain = census_income$increase_value)

leaflet(options = leafletOptions(zoomSnap = 0.1)) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    data = st_geometry(community),
    color = "black",
    fillColor = "gray",
    popup = neigh_labels,
    fillOpacity = .1,
    weight = 5,
    group = "Community Areas"
  ) %>% 
  addPolygons(
    data = st_geometry(census_income),
    color = "black",
    fillColor = median_increase_value(census_income$increase_value),
    fillOpacity = .6,
    weight = 1
  ) %>% 
  addLegend(
    title = "Increase in Home Value",
    pal = median_increase_value,
    values = census_income$increase_value,
    position = "topright"
  ) 

```
